### 2018-09-05 notebooks/poster_analysis.Rmd
added folder for notebooks

---
title: "Poster Data Analysis"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

# Description
This notebook presents the analysis of whole genome data for the poster "", presented at the ASM NGS Conference in Washington D.C. September 2018.

```{r, include = FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr)

# Functions

func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

## COnfidence intervals
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Function that returns a filtered dataframe based on the string "genes"
filter_mut_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes)
  return(df)
}

# Function that returns a filtered dataframe based on the string "acquired_genes"
filter_acquired_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in acquired_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene))
  return(df)
}

```

# Import data
This chunk imports data from MLST and ARIBA AMR gene detection results. Isolate data is also imported.
```{r}
# Isolate data, including DNA extraction results, species,
# method of isolation, and sequencing methodology
isolate_data <- read.table("../data/total_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

# ARIBA MLST results for each isolate
mlst_results <- read.table("../data/mlst/mlst_results.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

# ARIBA AMR gene results (megares database, for mutations in intrinsic genes)
megares_results <- read.table("../data/megares_results.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

# ARIBA AMR gene results (resfinder database, for acquired genes)
resfinder_results <- read.table("../data/resfinder_results.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

```

# Figures
## Percent presence of acquired genes in total
```{r}
resfinder_results %>%
  gather(gene, result, -c(saksnr, type)) %>%
  group_by(gene, result) %>%
  count() %>%
  ungroup() %>%
  mutate(result = if_else(result == 1, "Present", "Absent")) %>%
  spread(result, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1)) %>%
  ggplot(aes(gene, Percent))+
  geom_col(color = "black")+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6)+
  labs(y = "Percent (%) of isolates")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())
```

## Percent presence of mutations in intrinsic genes in total
```{r}
megares_results %>%
  gather(gene, result, -c(saksnr, type)) %>%
  group_by(gene, result) %>%
  count() %>%
  ungroup() %>%
  mutate(compl = complete.cases(.),
         result = if_else(result == 1, "Present", "Absent")) %>%
  filter(compl == TRUE) %>%
  select(-compl) %>%
  spread(result, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1)) %>%
  ggplot(aes(gene, Percent))+
  geom_col(color = "black")+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6)+
  labs(y = "Percent (%) of isolates")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())
```

## Acquired genes per animal species
```{r}

gene_names <- names(resfinder_results)

gene_names <- gene_names[-c(1,2)]

resfinder_results %>%
  left_join(isolate_data, by = "saksnr") %>%
  left_join(isolate_data, by = c("saksnr" = "plate_id")) %>%
  mutate(species.y = if_else(is.na(species.y) == TRUE, species.x, species.y),
         saksnr.y = if_else(is.na(saksnr.y) == TRUE, saksnr, saksnr.y),
         method.y = if_else(is.na(method.y) == TRUE, method.x, method.y),
         id = paste(saksnr.y, method.y, sep = "-"),
         dupl = duplicated(id)) %>%
  filter(dupl == FALSE) %>%
  gather(key, value, gene_names) %>%
  group_by(key, value, species.y) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1)) %>%
  ggplot(aes(key, Percent, fill = species.y))+
  geom_col(color = "black",
           position = position_dodge(1))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(1))+
  labs(y = "Percent (%) of isolates")+
  guides(fill = FALSE)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())+
  facet_wrap(~species.y)
```

# Mutations in intrinsic genes per animal species
```{r}
mut_names <- names(megares_results)

mut_names <- mut_names[-c(1,2)]

megares_results %>%
  left_join(isolate_data, by = "saksnr") %>%
  left_join(isolate_data, by = c("saksnr" = "plate_id")) %>%
  mutate(species.y = if_else(is.na(species.y) == TRUE, species.x, species.y),
         saksnr.y = if_else(is.na(saksnr.y) == TRUE, saksnr, saksnr.y),
         method.y = if_else(is.na(method.y) == TRUE, method.x, method.y),
         id = paste(saksnr.y, method.y, sep = "-"),
         dupl = duplicated(id)) %>%
  filter(dupl == FALSE) %>%
  gather(key, value, mut_names) %>%
  group_by(key, value, species.y) %>%
  count() %>%
  ungroup() %>%
  mutate(compl = complete.cases(.),
         value = if_else(value == 1, "Mutant", "Wild Type")) %>%
  filter(compl == TRUE) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Mutant + `Wild Type`,
         Percent = round(Mutant/Total*100, 1),
         lwr = round(get_binCI(Mutant, Total)[1], 1),
         upr = round(get_binCI(Mutant, Total)[2], 1)) %>%
  ggplot(aes(key, Percent, fill = species.y))+
  geom_col(color = "black",
           position = position_dodge(1))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(1))+
  labs(y = "Percent (%) of isolates")+
  guides(fill = FALSE)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())+
  facet_wrap(~species.y)
```

### 2018-09-05 notebooks/poster_analysis.Rmd
updated plots and started on MIC plots

@@
 isolate_data <- read.table("../data/total_data.txt",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
+
+# Total MIC profile for all isolates in this study
+mic_data <- read.table("../data/total_MIC.txt",
                            sep = "\t",
@@
 
-# Mutations in intrinsic genes per animal species
+## Mutations in intrinsic genes per animal species
 ```{r}
@@
 ```
+
+## MIC plots
+```{r}
+test <- megares_results %>%
+  left_join(isolate_data, by = "saksnr") %>%
+  left_join(isolate_data, by = c("saksnr" = "plate_id")) %>%
+  mutate(species.y = if_else(is.na(species.y) == TRUE, species.x, species.y),
+         saksnr.y = if_else(is.na(saksnr.y) == TRUE, saksnr, saksnr.y),
+         method.y = if_else(is.na(method.y) == TRUE, method.x, method.y),
+         NAL.y = if_else(is.na(NAL.y) == TRUE, NAL.x, NAL.y),
+         CIP.y = if_else(is.na(CIP.y) == TRUE, CIP.x, CIP.y),
+         id = paste(saksnr.y, method.y, sep = "-"),
+         dupl = duplicated(id)) %>%
+  filter(dupl == FALSE) %>%
+  gather(key, value, mut_names) %>%
+  group_by(key, value, CIP.y) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(compl = complete.cases(.),
+         value = if_else(value == 1, "Mutant", "Wild Type")) %>%
+  filter(compl == TRUE) %>%
+  spread(value, n, fill = 0)
+```

### 2018-09-06 notebooks/poster_analysis.Rmd
added poster title

@@
 # Description
-This notebook presents the analysis of whole genome data for the poster "", presented at the ASM NGS Conference in Washington D.C. September 2018.
+This notebook presents the analysis of whole genome data for the poster "Quinolone resistance mechanisms found in E. coli from four animal species in Norway
+", presented at the ASM NGS Conference in Washington D.C. September 2018.
 

### 2018-09-06 ariba_script.R
added fix for id's

@@ -28,15 +28,30 @@ mut_filtered <- filter_mut_table(mut_table)
 acquired_table <- create_acquired_table(clean_acquired_data)
 acquired_filtered <- filter_acquired_table(acquired_table)
 
+filtered_mut_names <- unique(mut_filtered$gene)
+filtered_acquired_names <- unique(acquired_filtered$gene)
+
 mut_complete <- mut_filtered %>%
   spread(gene, result) %>%
-  mutate(ref = gsub("(.*?)\\_.*", "\\1", ref)) %>%
-  rename("saksnr" = ref)
+  rename("plate_id" = ref) %>%
+  mutate(plate_id = gsub("(.*?)\\_.+", "\\1", plate_id),
+         plate_id = sub("^\\d*-?(\\d{4}-.*)", "\\1", plate_id),
+         plate_id = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", plate_id)) %>%
+  left_join(isolate_data2) %>%
+  mutate(dupl = duplicated(id)) %>%
+  filter(dupl == FALSE) %>%
+  select_at(.vars = vars(id, type, filtered_mut_names))
 
 acquired_complete <- acquired_filtered %>%
   spread(gene, result) %>%
-  mutate(ref = gsub("(.*?)\\_.*", "\\1", ref)) %>%
-  rename("saksnr" = ref)
+  rename("plate_id" = ref) %>%
+  mutate(plate_id = gsub("(.*?)\\_.+", "\\1", plate_id),
+         plate_id = sub("^\\d*-?(\\d{4}-.*)", "\\1", plate_id),
+         plate_id = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", plate_id)) %>%
+  left_join(isolate_data2) %>%
+  mutate(dupl = duplicated(id)) %>%
+  filter(dupl == FALSE) %>%
+  select_at(.vars = vars(id, type, filtered_acquired_names))
 
 write.table(mut_complete, "data/megares_results.txt", sep = "\t", row.names = F)
 write.table(acquired_complete, "data/resfinder_results.txt", sep = "\t", row.names = F)

### 2018-09-06 mlst_data.R
deleted

@@ -1,76 +1,0 @@
-total_df_offline <- read_excel("data/komplett_oversikt_isolater_art2.xlsx") %>%
-  rename(species = NORMart,
-         CIP = T_CIP,
-         NAL = U_NAL) %>%
-  mutate(species = case_when(species == "avi" ~ "Wild bird",
-                             species == "chi" ~ "Broiler",
-                             species == "fox" ~ "Red fox",
-                             species == "pig" ~ "Pig"),
-         species = factor(species),
-         method = factor(method),
-         qubit = as.numeric(qubit),
-         `260/280` = as.numeric(`260/280`),
-         `260/230` = as.numeric(`260/230`),
-         antres = as.numeric(antres),
-         read_size = factor(as.character(read_size)),
-         group = factor(group),
-         CIP = factor(CIP,
-                      levels = c("0.03",
-                                 "0.06",
-                                 "0.12",
-                                 "0.25",
-                                 "0.5",
-                                 "1",
-                                 "2",
-                                 "4",
-                                 "8",
-                                 "16"),
-                      ordered = TRUE),
-         NAL = factor(NAL,
-                      levels = c("4",
-                                 "8",
-                                 "16",
-                                 "32",
-                                 "64",
-                                 "128",
-                                 "256"),
-                      ordered = TRUE))
-
-mlst_data <- read.table("data/mlst/mlst_results.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE) %>%
-  mutate(plate_id = saksnr)
-
-
-
-isolate_data <- total_df_offline %>%
-  select(saksnr, plate_id, species, method)
-
-
-mlst_clean <- mlst_data %>%
-  left_join(isolate_data, by = "plate_id") %>%
-  left_join(isolate_data, by = c("saksnr.x" = "saksnr")) %>%
-  mutate(saksnr.y = if_else(is.na(saksnr.y) == TRUE, saksnr.x, saksnr.y),
-         species.x = if_else(is.na(species.x) == TRUE, species.y, species.x),
-         method.x = if_else(is.na(method.x) == TRUE, method.y, method.x)) %>%
-  select(saksnr.y, plate_id.x, species.x, method.x, ST, adk, fumC, gyrB, icd, mdh, purA, recA) %>%
-  rename("saksnr" = saksnr.y,
[…]

### 2018-09-06 notebooks/poster_analysis.Rmd
fixed imported files to correct ID settings

--- 2018-09-05 notebooks/poster_analysis.Rmd
@@
 # method of isolation, and sequencing methodology
-isolate_data <- read.table("../data/total_data.txt",
+isolate_data <- read.table("../data/isolate_data.txt",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
+
+# Total MIC profile for all isolates in this study
+mic_data <- read.table("../data/mic_data.txt",
                            sep = "\t",
@@
 # ARIBA MLST results for each isolate
-mlst_results <- read.table("../data/mlst/mlst_results.txt",
+mlst_results <- read.table("../data/mlst_results.txt",
                            sep = "\t",
@@
 
+total_data <- isolate_data %>%
+  left_join(mic_data) %>%
+  left_join(mlst_results) %>%
+  rename("m_gyrB" = gyrB) %>%
+  left_join(resfinder_results) %>%
+  select(-type) %>%
+  left_join(megares_results) %>%
+  select(-type)
 ```
@@
 resfinder_results %>%
-  gather(gene, result, -c(saksnr, type)) %>%
+  gather(gene, result, -c(id, type)) %>%
   group_by(gene, result) %>%
@@
 megares_results %>%
-  gather(gene, result, -c(saksnr, type)) %>%
+  gather(gene, result, -c(id, type)) %>%
   group_by(gene, result) %>%
@@
 ```{r}
-
 gene_names <- names(resfinder_results)
@@
 resfinder_results %>%
-  left_join(isolate_data, by = "saksnr") %>%
-  left_join(isolate_data, by = c("saksnr" = "plate_id")) %>%
-  mutate(species.y = if_else(is.na(species.y) == TRUE, species.x, species.y),
-         saksnr.y = if_else(is.na(saksnr.y) == TRUE, saksnr, saksnr.y),
-         method.y = if_else(is.na(method.y) == TRUE, method.x, method.y),
-         id = paste(saksnr.y, method.y, sep = "-"),
-         dupl = duplicated(id)) %>%
-  filter(dupl == FALSE) %>%
+  left_join(isolate_data, by = "id") %>%
   gather(key, value, gene_names) %>%
-  group_by(key, value, species.y) %>%
+  group_by(key, value, species) %>%
   count() %>%
@@
          upr = round(get_binCI(Present, Total)[2], 1)) %>%
-  ggplot(aes(key, Percent, fill = species.y))+
+  ggplot(aes(key, Percent, fill = species))+
   geom_col(color = "black",
@@
         axis.title.x = element_blank())+
-  facet_wrap(~species.y)
+  facet_wrap(~species)
 ```
 
-# Mutations in intrinsic genes per animal species
+## Mutations in intrinsic genes per animal species
 ```{r}
@@
 megares_results %>%
-  left_join(isolate_data, by = "saksnr") %>%
-  left_join(isolate_data, by = c("saksnr" = "plate_id")) %>%
-  mutate(species.y = if_else(is.na(species.y) == TRUE, species.x, species.y),
-         saksnr.y = if_else(is.na(saksnr.y) == TRUE, saksnr, saksnr.y),
-         method.y = if_else(is.na(method.y) == TRUE, method.x, method.y),
-         id = paste(saksnr.y, method.y, sep = "-"),
-         dupl = duplicated(id)) %>%
-  filter(dupl == FALSE) %>%
+  left_join(isolate_data, by = "id") %>%
   gather(key, value, mut_names) %>%
-  group_by(key, value, species.y) %>%
+  group_by(key, value, species) %>%
   count() %>%
   ungroup() %>%
-  mutate(compl = complete.cases(.),
-         value = if_else(value == 1, "Mutant", "Wild Type")) %>%
+  mutate(value = if_else(value == 1, "Mutant", "Wild Type"),
+         compl = complete.cases(.)) %>%
   filter(compl == TRUE) %>%
@@
          upr = round(get_binCI(Mutant, Total)[2], 1)) %>%
-  ggplot(aes(key, Percent, fill = species.y))+
+  ggplot(aes(key, Percent, fill = species))+
   geom_col(color = "black",
@@
         axis.title.x = element_blank())+
-  facet_wrap(~species.y)
+  facet_wrap(~species)
 ```
+
+## MIC plots
+```{r}
+
+```

### 2018-09-07 ariba_script.R
updated flag information and reporting

@@ -21,6 +21,10 @@ acquired_data <- get_ariba_data(resfinder_results)
 clean_mut_data <- fix_gene_names(mut_data)
 clean_acquired_data <- fix_gene_names(acquired_data)
 
+# Check flags
+mut_flags <- check_flags(clean_mut_data)
+acquired_flags <- check_flags(clean_acquired_data)
+
 # Wrangle
 mut_table <- create_mut_table(clean_mut_data)
 mut_filtered <- filter_mut_table(mut_table)

### 2018-09-07 src/ariba_functions.R
added functions for flag reporting

@@ -74,6 +74,18 @@ flag_selection <- c("19","27","147","155","403",
                     "411","915","923","787","795",
                     "531","539","659","667","787","795")  
 
+# Function that returns info on flag selection
+check_flags <- function(df) {
+  df <- df %>%
+    select(ref, gene_names, flag, ref_ctg_change) %>%
+    mutate(flag_result = flag %in% flag_selection,
+           flag_result = as.integer(flag_result),
+           ref = gsub("(.*?)\\_.+", "\\1", ref),
+           ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
+           ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref))
+  return(df)
+}
+
 # Function that handles megares data and returns a data frame with information on whether a gene
 # is mutated or not. Includes control for QRDR in gyrA.
 create_mut_table <- function(df) {
@@ -86,7 +98,7 @@ create_mut_table <- function(df) {
     summarise_all(funs(func_paste)) %>%
     mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
     select(-c(id, flag)) %>%
-    gather(gene, result,-ref) %>%
+    gather(gene, result, -ref) %>%
     mutate(gyrA_result = result %>% # control for mutation within QRDR for gyrA
              str_extract_all("\\d+") %>% # from reprex package
              map(as.integer) %>% # converts all to integer

### 2018-09-07 notebooks/poster_analysis.Rmd
updated figures and included heatmap and mic-plot

@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr)
+pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats)
 
@@
                                                 c("lwr", "upr")))
-
-# Function that returns a filtered dataframe based on the string "genes"
-filter_mut_table <- function(df) {
-  report_genes <- unique(df$gene)
-  grep_genes <- c()
-  for (gene in genes) {
-    for (g in report_genes) {
-      if (grepl(gene, g, ignore.case = T) == TRUE) {
-        grep_genes <- c(grep_genes, g)
-      }
-    }
-  }
-  df <- df %>%
-    filter(gene %in% grep_genes)
-  return(df)
-}
-
-# Function that returns a filtered dataframe based on the string "acquired_genes"
-filter_acquired_table <- function(df) {
-  report_genes <- unique(df$gene)
-  grep_genes <- c()
-  for (gene in acquired_genes) {
-    for (g in report_genes) {
-      if (grepl(gene, g, ignore.case = T) == TRUE) {
-        grep_genes <- c(grep_genes, g)
-      }
-    }
-  }
-  df <- df %>%
-    filter(gene %in% grep_genes) %>%
-    mutate(gene = gsub("_", "", gene))
-  return(df)
-}
-
 ```
@@
 total_data <- isolate_data %>%
-  left_join(mic_data) %>%
-  left_join(mlst_results) %>%
+  left_join(mic_data, by = "id") %>%
+  left_join(mlst_results, by = "id") %>%
   rename("m_gyrB" = gyrB) %>%
-  left_join(resfinder_results) %>%
+  left_join(resfinder_results, by = "id") %>%
   select(-type) %>%
-  left_join(megares_results) %>%
+  left_join(megares_results, by = "id") %>%
   select(-type)
@@
                 position = position_dodge(1))+
+  scale_fill_brewer(palette = "Paired")+
   labs(y = "Percent (%) of isolates")+
@@
 
-## MIC plots
-```{r}
+## MIC plot
 
+```{r message=FALSE, warning=FALSE}
+selected_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")
+selected_mut <- c("gyrA","gyrB","parC","parE","marR")
+
+micplot_data <- mic_data %>%
+  select(id, CIP) %>%
+  left_join(megares_results, by = "id") %>%
+  left_join(resfinder_results, by = "id") %>%
+  select(-c(acrB, type.x, type.y)) %>%
+  mutate(compl = complete.cases(.)) %>%
+  filter(compl == TRUE) %>%
+  select_at(vars(id, CIP,selected_mut, selected_genes)) %>%
+  mutate(none = if_else(rowSums(.[,3:ncol(.)]) != 0, 0, 1)) %>%
+  group_by_at(vars(selected_mut, selected_genes, none)) %>%
+  { mutate(ungroup(.), group = group_indices(.)) } %>%
+  gather(gene, value, -c(id, CIP, group)) %>%
+  mutate(gene = ifelse(value == 0 & gene != "none", NA, gene),
+         value = ifelse(gene == "none" & value == 0, NA, value)) %>%
+  na.omit()
+
+factor_breaks <- c(0.015, 0.03, 0.06, 0.12, 0.25, 0.5, 1, 2, 4, 8, 16)
+
+factor_levels <- c("0.015", "0.03", "0.06", "0.12", "0.25", "0.5", "1", "2", "4", 
+                   "8", "16")
+
+p0 <- micplot_data %>%
+  group_by(id, group) %>%
+  count() %>%
+  group_by(group) %>%
+  count() %>%
+  ggplot(aes(factor(group), nn, fill = factor(group)))+
+  geom_col(color = "black")+
+  guides(fill = FALSE)+
+  scale_fill_viridis(discrete = TRUE,
+                     direction = -1)+
+  theme_light()+
+  theme(axis.title = element_blank(),
+        axis.text.y = element_text(size = 8),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid = element_blank())
+
+p1 <- micplot_data %>%
+  group_by(CIP, group, gene) %>%
+  count() %>%
+  ggplot(aes(factor(group), CIP, fill = factor(group), size = n))+
+  geom_point(pch = 21)+
+  scale_y_continuous(trans = "log10",
+                     labels = factor_levels,
+                     breaks = factor_breaks)+
+  guides(fill = FALSE,
+         size = FALSE)+
+  scale_fill_viridis(discrete = TRUE,
+                     direction = -1)+
+  theme_light()+
+  theme(axis.title = element_blank(),
+        axis.text.y = element_text(size = 8),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid = element_blank())
+
+p2 <- ggplot(micplot_data,
+       aes(factor(group),
+           factor(gene),
+           fill = factor(group)))+
+  geom_point(shape = 23,
+             pch = 21,
+             size = 2)+
+  geom_hline(yintercept = 5.5)+
+  guides(fill = FALSE)+
+  scale_fill_viridis(discrete = TRUE,
+                     direction = -1)+
+  labs(y = "Genes")+
+  theme_light()+
+  theme(axis.title = element_blank(),
+        axis.text.x = element_blank(),
+        axis.text.y = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        panel.grid = element_blank())
+
+plot_grid(p0, p1, p2, align = "v", nrow = 3, rel_heights = c(1/4, 2/4, 1/4))
 ```
+
+## Heatmap plot
+
+```{r, fig.width = 10, fig.height=30}
+species_data <- select(isolate_data, c(id, species))
+
+megares_long <- megares_results %>%
+  gather(gene, value, -c(id, type)) %>%
+  filter(is.na(value) == FALSE)
+
+resfinder_long <- resfinder_results %>%
+  gather(gene, value, -c(id, type))
+
+megares_long %>%
+  rbind(resfinder_long) %>%
+  mutate(type2 = ifelse(type == "mut", 1, 2)) %>%
+  spread(id, value, fill = 0) %>%
+  gather(key = "id", value = "result", -c(gene, type, type2)) %>%
+  left_join(species_data, by = "id") %>%
+  ggplot(aes(fct_reorder(gene, type2),
+             factor(id,
+                    levels = unique(.[order(.$species,
+                                                          .$id),
+                                                    "id",
+                                                    drop = TRUE])),
+             fill = species,
+             alpha = result)) +
+  geom_tile(color = "white")+
+  geom_vline(xintercept = 14.5)+
+  scale_fill_brewer(palette = "Paired")+
+  theme_minimal()+
+  labs(fill = NULL)+
+  theme(axis.text.x = element_text(angle = 90,
+                                   hjust = 1,
+                                   vjust = 0.3,
+                                   size = 7),
+        axis.title = element_blank(),
+        axis.text.y = element_blank(),
+        panel.grid = element_blank(),
+        legend.position = "right")+
+  guides(alpha = FALSE)+
+  coord_fixed()
+```

### 2018-09-08 notebooks/poster_analysis.Rmd
updated heatmap to include NA values

@@
 megares_long <- megares_results %>%
-  gather(gene, value, -c(id, type)) %>%
-  filter(is.na(value) == FALSE)
+  gather(gene, value, -c(id, type))
 
@@
   rbind(resfinder_long) %>%
-  mutate(type2 = ifelse(type == "mut", 1, 2)) %>%
+  mutate(type2 = ifelse(type == "mut", 1, 2),
+         value = factor(ifelse(is.na(value) == TRUE, 2, value))) %>%
   spread(id, value, fill = 0) %>%
@@
   left_join(species_data, by = "id") %>%
+  mutate(result = ifelse(result == 2, NA, result),
+         fillcol = paste(species, result, sep = "_"),
+         fillcol = ifelse(is.na(result) == TRUE, NA, fillcol)) %>%
   ggplot(aes(fct_reorder(gene, type2),
              factor(id,
-                    levels = unique(.[order(.$species,
-                                                          .$id),
-                                                    "id",
-                                                    drop = TRUE])),
-             fill = species,
+                    levels = unique(.[order(.$species,.$id),"id",drop = TRUE])),
+             fill = fillcol,
              alpha = result)) +
@@
   geom_vline(xintercept = 14.5)+
-  scale_fill_brewer(palette = "Paired")+
+  scale_fill_brewer(palette = "Paired",
+                    na.value = "#6b6767")+
   theme_minimal()+
@@
   coord_fixed()
+
 ```

### 2018-09-08 notebooks/poster_analysis.Rmd
added tables for sample information

@@
 
+species_data <- select(isolate_data, c(id, species))
+
 total_data <- isolate_data %>%
@@
   select(-type)
+```
+
+# Tables
+
+## Samples per species
+```{r}
+per_species <- isolate_data %>%
+  group_by(species) %>%
+  count()
+per_species
+```
+
+## Samples per sequencing method
+```{r}
+per_seq <- isolate_data %>%
+  group_by(sequencer, libprep, read_length) %>%
+  count()
+per_seq
+```
+
+## MLST Data
+
+### Total MLST 
+```{r}
+total_mlst <- mlst_results %>%
+  left_join(species_data, by = "id") %>%
+  group_by(ST) %>%
+  count()
+total_mlst
+```
+
+### MLST per animal species
+```{r}
+mlst_per_species <- mlst_results %>%
+  left_join(species_data, by = "id") %>%
+  group_by(ST, species) %>%
+  count()
+mlst_per_species
 ```
@@
 
-```{r, fig.width = 10, fig.height=30}
-species_data <- select(isolate_data, c(id, species))
-
+```{r fig.height=30, fig.width=10, message=FALSE, warning=FALSE}
 megares_long <- megares_results %>%

### 2018-09-09 notebooks/poster_analysis.Rmd
added correlation plot for MIC's

@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats)
+pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data)
 
@@
 
-## MIC plot
+## MIC plots
 
+### MIC Distribution for generated groups
 ```{r message=FALSE, warning=FALSE}
@@
 
+### MIC correlation plots per animal species
+
+```{r}
+test <- mic_data %>%
+  select(-antres) %>%
+  left_join(species_data, by = "id") %>%
+  mutate(SMX_R = ifelse(SMX > 64, 1, 0),
+         TMP_R = ifelse(TMP > 2, 1, 0),
+         CIP_R = ifelse(CIP > 0.06, 1, 0),
+         TET_R = ifelse(TET > 8, 1, 0),
+         NAL_R = ifelse(NAL > 16, 1, 0),
+         CTX_R = ifelse(CTX > 0.25, 1, 0),
+         CHL_R = ifelse(CHL > 16, 1, 0),
+         AMP_R = ifelse(AMP > 8, 1, 0),
+         GEN_R = ifelse(GEN > 2, 1, 0)) %>%
+  select(contains("_R"), species) %>%
+  split(., f = .$species) %>%
+  lapply(., function(x) select(x, -species))
+
+create_corr_matrix <- function(df) {
+  df <- df %>%
+    as.matrix %>%
+    cor %>%
+    as.data.frame %>%
+    rownames_to_column(var = "var1") %>%
+    gather(var2, value, -var1) %>%
+    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
+           var2 = gsub("(.*?)_R$", "\\1", var2))
+  return(df)
+}
+
+corr_matrices <- lapply(test, function(x) create_corr_matrix(x)) %>%
+  bind_rows(.id = "species")
+
+cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a", "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
+
+ggplot(corr_matrices, aes(var1, var2, fill = value))+
+  geom_tile(color = "white")+
+  scale_fill_gradientn(colors=cols,
+                       na.value="grey95",
+                       limits=c(-1, 1),
+                       guide=guide_colourbar(ticks=T,nbin=10,
+                                             barheight=.5,label=T, 
+                                             barwidth=10))+
+  labs(fill = NULL)+
+  theme_classic()+
+  theme(axis.text = element_text(size = 7),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
+        axis.title = element_blank(),
+        legend.position="bottom",
+        legend.justification="center",
+        legend.direction="horizontal")+
+  coord_fixed()+
+  facet_wrap(~species)
+```
+
 ## Heatmap plot

### 2018-09-09 notebooks/poster_analysis.Rmd
added statistical test for comparison of qnr-genes

@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data)
+pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr)
 
@@
 ## Acquired genes per animal species
+### Total acquired genes per animal species
 ```{r}
@@
   facet_wrap(~species)
+```
+
+### Qnr-genes per animal species
+```{r}
+qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")
+
+qnr_data <- resfinder_results %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, qnr_genes) %>%
+  group_by(key, value, species) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total*100, 1),
+         lwr = round(get_binCI(Present, Total)[1], 1),
+         upr = round(get_binCI(Present, Total)[2], 1))
+
+ggplot(qnr_data, aes(key, Percent, fill = species))+
+  geom_col(color = "black",
+           position = position_dodge(0.9))+
+  geom_errorbar(aes(ymin = lwr, ymax = upr),
+                width = 0.6,
+                position = position_dodge(0.9))+
+  scale_fill_brewer(palette = "Paired")+
+  labs(y = "Percent (%) of isolates",
+       fill = NULL)+
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
+        axis.title.x = element_blank())
 ```
@@
 
-cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a", "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
+cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
+          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
 
@@
 ```
+
+# Statistics
+
+## Significance test for the percent presence of qnrB19 and qnrS1 between animal species
+```{r message=FALSE, warning=FALSE}
+specific_qnr <- qnr_data %>%
+  filter(key %in% c("qnrB19", "qnrS1")) %>%
+  select(key, species, Present, Absent) %>%
+  split(f = .$key)
+
+do_chisq_test <- function(df) {
+  df <- df %>%
+    slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
+          filter(Var2 < Var1) %>% t) %>%
+    mutate(test = rep(1:(n()/2), each = 2)) %>%
+    group_by(test) %>%
+    do(data_frame(gene = .$key,
+                  test = first(.$test),
+                  species1 = first(.$species),
+                  species2 = last(.$species),
+                  data = list(matrix(c(.$Absent,
+                                       .$Present),
+                                     ncol = 2)))) %>%
+    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
+    mutate(p.value = map_dbl(chi_test, "p.value"),
+           dupl = duplicated(test)) %>%
+    filter(dupl == FALSE) %>%
+    select(gene, species1, species2, p.value) %>%
+    filter(p.value < 0.05)
+  return(df)
+}
+
+p_values <- lapply(specific_qnr, function(x) do_chisq_test(x)) %>%
+  bind_rows(.id = "gene")
+
+p_values
+```

### 2018-09-09 notebooks/poster_analysis.Rmd
changed output of stats table
added chisq tests for all acquired and all mutations

@@
 
+```
+
+# Functions
+This chunk lists the functions used in this notebook.
+```{r}
 # Functions
@@
                                                 c("lwr", "upr")))
+
+# Creates a correlation matrix from 1/0 data from MIC values
+create_corr_matrix <- function(df) {
+  df <- df %>%
+    as.matrix %>%
+    cor %>%
+    as.data.frame %>%
+    rownames_to_column(var = "var1") %>%
+    gather(var2, value, -var1) %>%
+    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
+           var2 = gsub("(.*?)_R$", "\\1", var2))
+  return(df)
+}
+
+# Creates a data frame with p-values for comparing presence of specific genes between animal species
+do_chisq_test <- function(df) {
+  df <- df %>%
+    slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
+          filter(Var2 < Var1) %>% t) %>%
+    mutate(test = rep(1:(n()/2), each = 2)) %>%
+    group_by(test) %>%
+    do(data_frame(gene = .$key,
+                  test = first(.$test),
+                  species1 = first(.$species),
+                  species2 = last(.$species),
+                  data = list(matrix(c(.$Absent,
+                                       .$Present),
+                                     ncol = 2)))) %>%
+    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
+    mutate(p.value = map_dbl(chi_test, "p.value"),
+           dupl = duplicated(test)) %>%
+    filter(dupl == FALSE) %>%
+    select(gene, species1, species2, p.value) %>%
+    filter(p.value < 0.05)
+  return(df)
+}
 ```
@@
 
-create_corr_matrix <- function(df) {
-  df <- df %>%
-    as.matrix %>%
-    cor %>%
-    as.data.frame %>%
-    rownames_to_column(var = "var1") %>%
-    gather(var2, value, -var1) %>%
-    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
-           var2 = gsub("(.*?)_R$", "\\1", var2))
-  return(df)
-}
-
 corr_matrices <- lapply(test, function(x) create_corr_matrix(x)) %>%
@@
 
-## Significance test for the percent presence of qnrB19 and qnrS1 between animal species
+## Significance test for the percent presence of acquired genes between animal species
+Only significant values will be presented in the table below (p < 0.05).
 ```{r message=FALSE, warning=FALSE}
-specific_qnr <- qnr_data %>%
-  filter(key %in% c("qnrB19", "qnrS1")) %>%
+acquired_genes_list <- resfinder_results %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, gene_names) %>%
+  group_by(key, value, species) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  rowwise() %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total*100, 1),
+         lwr = round(get_binCI(Present, Total)[1], 1),
+         upr = round(get_binCI(Present, Total)[2], 1)) %>%
   select(key, species, Present, Absent) %>%
@@
 
-do_chisq_test <- function(df) {
-  df <- df %>%
-    slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
-          filter(Var2 < Var1) %>% t) %>%
-    mutate(test = rep(1:(n()/2), each = 2)) %>%
-    group_by(test) %>%
-    do(data_frame(gene = .$key,
-                  test = first(.$test),
-                  species1 = first(.$species),
-                  species2 = last(.$species),
-                  data = list(matrix(c(.$Absent,
-                                       .$Present),
-                                     ncol = 2)))) %>%
-    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
-    mutate(p.value = map_dbl(chi_test, "p.value"),
-           dupl = duplicated(test)) %>%
-    filter(dupl == FALSE) %>%
-    select(gene, species1, species2, p.value) %>%
-    filter(p.value < 0.05)
-  return(df)
-}
+p_values_acquired <- lapply(acquired_genes_list, function(x) do_chisq_test(x)) %>%
+  bind_rows()
 
-p_values <- lapply(specific_qnr, function(x) do_chisq_test(x)) %>%
-  bind_rows(.id = "gene")
+p_values_acquired
+```
 
-p_values
+## Significance test for the percent presence of mutations in intrinsic genes between animal species
+Only significant values will be presented in the table below (p < 0.05).
+
+```{r message=FALSE, warning=FALSE}
+mut_genes_list <- megares_results %>%
+  left_join(isolate_data, by = "id") %>%
+  gather(key, value, mut_names) %>%
+  filter(key != "acrB") %>%
+  group_by(key, value, species) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+  spread(value, n, fill = 0) %>%
+  select(key, species, Present, Absent) %>%
+  split(f = .$key)
+
+p_values_mut <-  lapply(mut_genes_list, function(x) do_chisq_test(x)) %>%
+  bind_rows()
+
+p_values_mut
 ```

### 2018-09-09 notebooks/poster_analysis.Rmd
minor updates

@@
   spread(value, n, fill = 0) %>%
-  rowwise() %>%
-  mutate(Total = Present + Absent,
-         Percent = round(Present/Total*100, 1),
-         lwr = round(get_binCI(Present, Total)[1], 1),
-         upr = round(get_binCI(Present, Total)[2], 1)) %>%
   select(key, species, Present, Absent) %>%

### 2018-09-10 notebooks/poster_analysis.Rmd
added correlation plots for genes

@@
 
+## Correlation plot for mutations in intrinsic genes
+```{r}
+corr_megares <- megares_results %>%
+  select(-c(type, acrB)) %>%
+  left_join(species_data, by = "id") %>%
+  mutate(compl = complete.cases(.)) %>%
+  filter(compl == TRUE) %>%
+  split(., f = .$species) %>%
+  lapply(., function(x) select(x, -c(species, id, compl)))
+
+mut_corr_matrices <- lapply(corr_megares, function(x) create_corr_matrix(x)) %>%
+  bind_rows(.id = "species")
+
+cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
+          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
+
+ggplot(mut_corr_matrices, aes(var1, var2, fill = value))+
+  geom_tile(color = "white")+
+  scale_fill_gradientn(colors=cols,
+                       na.value="grey95",
+                       limits=c(-1, 1),
+                       guide=guide_colourbar(ticks=T,nbin=10,
+                                             barheight=.5,label=T, 
+                                             barwidth=10))+
+  labs(fill = NULL)+
+  theme_classic()+
+  theme(axis.text = element_text(size = 7),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
+        axis.title = element_blank(),
+        legend.position="bottom",
+        legend.justification="center",
+        legend.direction="horizontal")+
+  coord_fixed()+
+  facet_wrap(~species)
+```
+
+## Correlation plot for acquired genes
+```{r message=FALSE, warning=FALSE}
+corr_resfinder <- resfinder_results %>%
+  select(-type) %>%
+  left_join(species_data, by = "id") %>%
+  split(., f = .$species) %>%
+  lapply(., function(x) select(x, -c(species, id)))
+
+acquired_corr_matrices <- lapply(corr_resfinder, function(x) create_corr_matrix(x)) %>%
+  bind_rows(.id = "species")
+
+cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
+          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")
+
+ggplot(acquired_corr_matrices, aes(var1, var2, fill = value))+
+  geom_tile(color = "white")+
+  scale_fill_gradientn(colors=cols,
+                       na.value="grey95",
+                       limits=c(-1, 1),
+                       guide=guide_colourbar(ticks=T,nbin=10,
+                                             barheight=.5,label=T, 
+                                             barwidth=10))+
+  labs(fill = NULL)+
+  theme_classic()+
+  theme(axis.text = element_text(size = 6),
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
+        axis.title = element_blank(),
+        legend.position="bottom",
+        legend.justification="center",
+        legend.direction="horizontal")+
+  coord_fixed()+
+  facet_wrap(~species)
+```
+
 ## MIC plots

### 2018-09-10 notebooks/poster_analysis.Rmd
minor fixes and added table of mlst profiles per species

@@
 
+## Total percentage of isolates with detected acquired genes and mutant intrinsic genes
+
+### Percent acquired genes
+```{r}
+resfinder_results %>%
+  select(-c(id, type)) %>%
+  mutate(total = rowSums(.),
+         result = if_else(total != 0, 1, 0)) %>%
+  group_by(result) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
+  spread(result, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2))
+
+```
+
+### Percent mutations in intrinsic genes
+```{r}
+megares_results %>%
+  select(-c(id, type, acrB)) %>%
+  mutate(total = rowSums(., na.rm = TRUE),
+         result = if_else(total != 0, 1, 0)) %>%
+  group_by(result) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
+  spread(result, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2))
+```
+
 ## MLST Data
@@
 mlst_per_species
+```
+
+### Number of sequence types per animal species
+```{r}
+total_ST_per_species <- mlst_per_species %>%
+  group_by(species) %>%
+  count()
+total_ST_per_species
 ```
@@
 ```{r}
-test <- mic_data %>%
+mic_per_species <- mic_data %>%
   select(-antres) %>%
@@
 
-corr_matrices <- lapply(test, function(x) create_corr_matrix(x)) %>%
+corr_matrices <- lapply(mic_per_species, function(x) create_corr_matrix(x)) %>%
   bind_rows(.id = "species")

### 2018-09-11 notebooks/poster_analysis.Rmd
removed acrB from analysis

@@
                            header = T,
-                           stringsAsFactors = F)
+                           stringsAsFactors = F) %>%
+  select(-acrB)
 
@@
 species_data <- select(isolate_data, c(id, species))
-
-total_data <- isolate_data %>%
-  left_join(mic_data, by = "id") %>%
-  left_join(mlst_results, by = "id") %>%
-  rename("m_gyrB" = gyrB) %>%
-  left_join(resfinder_results, by = "id") %>%
-  select(-type) %>%
-  left_join(megares_results, by = "id") %>%
-  select(-type)
 ```
@@
 megares_results %>%
-  select(-c(id, type, acrB)) %>%
+  select(-c(id, type)) %>%
   mutate(total = rowSums(., na.rm = TRUE),
@@
 corr_megares <- megares_results %>%
-  select(-c(type, acrB)) %>%
+  select(-type) %>%
   left_join(species_data, by = "id") %>%
@@
   left_join(resfinder_results, by = "id") %>%
-  select(-c(acrB, type.x, type.y)) %>%
+  select(-c(type.x, type.y)) %>%
   mutate(compl = complete.cases(.)) %>%
@@
   gather(key, value, mut_names) %>%
-  filter(key != "acrB") %>%
   group_by(key, value, species) %>%

### 2018-09-11 notebooks/poster_analysis.Rmd
updated figures and tables

@@
 
-```{r, include = FALSE}
+Please press the "code" button for each chunk to view code.
+```{r message=FALSE, warning=FALSE}
 # Libraries
@@
 pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr)
-
 ```
@@
 }
+
+"%not_in%" <- Negate("%in%")
+
 ```
@@
 # Import data
-This chunk imports data from MLST and ARIBA AMR gene detection results. Isolate data is also imported.
+This chunk imports data from MLST and ARIBA[1] AMR gene detection results. Isolate data is also imported.
+Four isolates were excluded due to bad sequence quality.
 ```{r}
@@
                            header = T,
-                           stringsAsFactors = F)
+                           stringsAsFactors = F) %>%
+  filter(id %not_in% c("2016-02-428-2-S","2016-02-486-2-S","2016-02-732-2-S","2016-17-565-1-S"))
 
@@
                            stringsAsFactors = F) %>%
-  select(-acrB)
+  select(-acrB) %>%
+  filter(id %not_in% c("2016-02-428-2-S","2016-02-486-2-S","2016-02-732-2-S","2016-17-565-1-S"))
 
@@
 # Tables
-
 ## Samples per species
@@
 ## Samples per sequencing method
+The isolates were sequenced by three different methods due to being part of different projects. The table below lists each sequencer used, library prep, and the resulting read length.
 ```{r}
@@
   group_by(sequencer, libprep, read_length) %>%
-  count()
+  count() %>%
+  rename("Sequencer" = sequencer,
+         "Library Prep" = libprep,
+         "Read Length" = read_length)
 per_seq
@@
 ### Percent acquired genes
+The total percentage of presence of acquired genes identified with ARIBA.
 ```{r}
@@
          lwr = round(get_binCI(Present, Total)[1], 2),
-         upr = round(get_binCI(Present, Total)[2], 2))
-
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr))
 ```
@@
 ### Percent mutations in intrinsic genes
+The total percentage of mutations in intrinsic genes identified with ARIBA.
 ```{r}
@@
          lwr = round(get_binCI(Present, Total)[1], 2),
-         upr = round(get_binCI(Present, Total)[2], 2))
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr))
 ```
@@
 ## MLST Data
-
 ### Total MLST
+The total amount of sequence types identified.
 ```{r}
@@
 ### MLST per animal species
+Sequence types for each animal species.
 ```{r}
@@
 ### Number of sequence types per animal species
+The total amount of unique sequence types per animal species.
 ```{r}
@@
 # Figures
+## Total MLST
+This figure presents the amount of isolates of each sequence type in total.
+```{r}
+mlst_results %>%
+  left_join(species_data, by = "id") %>%
+  group_by(ST) %>%
+  count() %>%
+  ggplot(aes(reorder(ST, -n), n))+
+  geom_col(color = "black")+
+  labs(x = "Sequence Types")+
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 90, size = 6, hjust = 1, vjust = 0.3))
+```
+
 ## Percent presence of acquired genes in total
+This plot presents the total percent occurrence of acquired genes identified with ARIBA, with 95 % confidence intervals.
 ```{r}
@@
 ## Percent presence of mutations in intrinsic genes in total
+This plot presents the total percent occurrence of mutations identified with ARIBA in intrinsic genes, with 95 % confidence intervals.
 ```{r}
@@
 ### Total acquired genes per animal species
+This plot presents the percent occurrence of acquired genes identified with ARIBA, with 95 % confidence intervals.
+Please note that not only quinolone resistance mechanisms is included in this plot.
+Colors: ColorBrewer[2]
 ```{r}
@@
 ### Qnr-genes per animal species
+This plot presents the percent occurrence of qnr-genes identified with ARIBA in the four animal species, with 95 % confidence intervals.
 ```{r}
@@
 ## Mutations in intrinsic genes per animal species
+This plot presents the percent occurrence of mutated intrinsic genes identified with ARIBA per animal species, with 95 % confidence intervals.
 ```{r}
@@
 
-## Correlation plot for mutations in intrinsic genes
+## Correlation plots
+These plots present the $\phi$-coefficient for the association between the presence/absence of mutations in intrinsic genes, and the presence/absence of acquired genes.
+### Mutations in intrinsic genes
 ```{r}
@@
 
-## Correlation plot for acquired genes
+### Presence/absence of acquired genes
 ```{r message=FALSE, warning=FALSE}
@@
 ## MIC plots
-
 ### MIC Distribution for generated groups
+This plot represents the MIC distribution for each auto-generated group of acquired genes and mutations combination.
+The plot on the bottom represents the group on the x-axis, and the acquired genes (top part of the horizontal line) and mutations in intrinsic genes (bottom part) that is present in the given group.
+The middle plot represents the MIC-distribution for the given group, where the size of the dot represents the amount of isolates at that MIC-value.
+Finally, the top plot represents the amount of isolates in the given group.
+All three plots have the same x-axis.
 ```{r message=FALSE, warning=FALSE}
@@
 ### MIC correlation plots per animal species
-
+This plot presents the $\phi$-coefficient for correlation between 1/0 values created from the MIC-values (1 = resistant, 0 = nonresistant).
 ```{r}
@@
 
+## Minimum Spanning Tree of MLST profiles
+Minimum spanning tree of sequence types, created in Phyloviz[3] with goeBURST distance (TLV).
+Light blue = Broiler, dark blue = Pig, light green = Red fox, and dark green = Wild bird.
+Distances between sequence types are denotd as numbers on the lines.
+![](../figures/mlst_tree_full.png)
+
 ## Heatmap plot
-
+This plot represents the individual resistance mechanisms (x-axis) present in each isolate (y-axis), for each animal species (colors).
+Grey squares represent genes/mutations that ARIBA flagged as bad quality in their report.
+Dark color indicates presence, while the light color indicates absence.
 ```{r fig.height=30, fig.width=10, message=FALSE, warning=FALSE}
@@
 Only significant values will be presented in the table below (p < 0.05).
-
 ```{r message=FALSE, warning=FALSE}
@@
 ```
+
+# References
+[1] ARIBA: rapid antimicrobial resistance genotyping directly from sequencing reads Hunt M, Mather AE, Sánchez-Busó L, Page AJ, Parkhill J , Keane JA, Harris SR.
+Microbial Genomics 2017.
+doi:  https://doi.org/110.1099/mgen.0.000131
+
+[2] ColorBrewer: An online tool for selecting color schemes for maps. The Cartographic Journal 40(1): 27-37
+
+[3] PHYLOViZ: phylogenetic inference and data visualization for sequence based typing methods, Francisco A. P. et al.,  BMC Bioinformatics 2012, 13:87. doi: https://doi.org/10.1186/1471-2105-13-87

### 2018-09-12 notebooks/poster_analysis.Rmd
added dendrogram for heatmap data

@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr)
+pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree)
 ```
@@
 ```{r fig.height=30, fig.width=10, message=FALSE, warning=FALSE}
+# Adjust MLST data frame
+sequence_types_adjusted <- mlst_results %>%
+  mutate_all(funs(gsub("\\*", "", .))) %>%
+  select(-ST)
+
+row.names(sequence_types_adjusted) <- sequence_types_adjusted$id
+
+mlst <- sequence_types_adjusted[,-1]
+
+# Calculate distance matrix from sequence typing alleles
+d.mlst.distances <- matrix(0, ncol=nrow(mlst), nrow=nrow(mlst))
+for (i in 1:(nrow(mlst)-1)) {
+  for (j in (i+1):nrow(mlst)){
+    d.mlst.distances[i,j] <- sum(mlst[i, ] != mlst[j, ])
+    d.mlst.distances[j,i] <- sum(mlst[i, ] != mlst[j, ])
+  }
+}
+
+# Create tree from distance matrix
+tree <- upgma(d.mlst.distances)
+id_label_tree <- tree 
+id_label_tree$tip.label <- sequence_types_adjusted$id
+tree$tip.label <- mlst_results$ST
+
+# Create dummy plot of id_label_tree to extract correct order of ID's
+p0 <- ggtree(id_label_tree)+
+  geom_treescale()+
+  geom_tiplab(align = TRUE, linesize = 0, hjust = -0.2, size = 1.5)+
+  xlim(0,3.6)
+
+# Get plot build data
+p1b <- ggplot_build(p0)
+
+# Save plot of tree
+p1 <- ggtree(tree)+
+  geom_treescale()+
+  geom_tiplab(align = TRUE, linesize = 0, hjust = -0.2, size = 1.5)+
+  xlim(0,3.6)
+
+# Create heatmap data frame from gene data
 megares_long <- megares_results %>%
@@
 
-megares_long %>%
+heatmap_data <- megares_long %>%
   rbind(resfinder_long) %>%
@@
          fillcol = paste(species, result, sep = "_"),
-         fillcol = ifelse(is.na(result) == TRUE, NA, fillcol)) %>%
-  ggplot(aes(fct_reorder(gene, type2),
-             factor(id,
-                    levels = unique(.[order(.$species,.$id),"id",drop = TRUE])),
+         fillcol = ifelse(is.na(result) == TRUE, NA, fillcol))
+
+# Arrange id-column after tip-labels in tree
+heatmap_data <- heatmap_data %>%
+  mutate(id = factor(id, levels = p1b$data[[5]] %>% arrange(y) %>% pull(label)))
+
+# Plot the heatmap
+p2 <- ggplot(heatmap_data, aes(fct_reorder(gene, type2),
+             factor(id),
              fill = fillcol,
@@
   geom_tile(color = "white")+
-  geom_vline(xintercept = 14.5)+
+  geom_vline(xintercept = 13.5)+
   scale_fill_brewer(palette = "Paired",
@@
   coord_fixed()
+
+# Arrange both plots into one figure
+grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(68,-10,149,0)),
+                         heights = c(1.9,-0.097)),
+             arrangeGrob(p2 + theme(axis.text.y=element_blank(),
+                                    plot.margin = margin(0,0,0,-8)),
+                         heights=c(0.98,-0.03)), 
+             ncol=2)
 

### 2018-09-12 notebooks/poster_analysis.Rmd
updated dendrogram

@@
 ## Heatmap plot
-This plot represents the individual resistance mechanisms (x-axis) present in each isolate (y-axis), for each animal species (colors).
+This plot represents the individual resistance mechanisms (x-axis) present in each isolate (y-axis), for each animal species (colors) and sequence type (dendrogram).
 Grey squares represent genes/mutations that ARIBA flagged as bad quality in their report.
 Dark color indicates presence, while the light color indicates absence.
+Part of the code below is taken from https://gist.github.com/scwatts/8c32d8f6dc44e352d89ae3de5f3a1125.
 ```{r fig.height=30, fig.width=10, message=FALSE, warning=FALSE}

### 2018-09-12 notebooks/poster_analysis.Rmd
updated color scheme on heatmap plot

@@
 ## Minimum Spanning Tree of MLST profiles
-Minimum spanning tree of sequence types, created in Phyloviz[3] with goeBURST distance (TLV).
+Minimum spanning tree of sequence types, created in Phyloviz[3] with goeBURST distance.
 Light blue = Broiler, dark blue = Pig, light green = Red fox, and dark green = Wild bird.
-Distances between sequence types are denotd as numbers on the lines.
+Distances between sequence types are denoted as numbers on the lines.
 ![](../figures/mlst_tree_full.png)
@@
 
+# Color vector
+palette <- c("Broiler_1" = "#a6cee3", "Broiler_0" = "#a6cee326",
+             "Pig_1" = "#1f78b4", "Pig_0" = "#1f78b426",
+             "Red fox_1" = "#b2df8a", "Red fox_0" = "#b2df8a26",
+             "Wild bird_1" = "#33a02c", "Wild bird_0" = "#33a02c26")
+
 # Plot the heatmap
@@
              factor(id),
-             fill = fillcol,
-             alpha = result)) +
+             fill = fillcol)) +
   geom_tile(color = "white")+
   geom_vline(xintercept = 13.5)+
-  scale_fill_brewer(palette = "Paired",
+  scale_fill_manual(values = palette,
                     na.value = "#6b6767")+
@@
         legend.position = "right")+
-  guides(alpha = FALSE)+
+  guides(fill = guide_legend(reverse = TRUE))+
   coord_fixed()

### 2018-09-12 notebooks/poster_analysis.Rmd
updated sizing of heatmap plot

@@
 # Arrange both plots into one figure
-grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(68,-10,149,0)),
+grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(60,-3,140,0)),
                          heights = c(1.9,-0.097)),

### 2018-09-12 notebooks/poster_analysis.Rmd
changed sizing of dendrogram

@@
 # Arrange both plots into one figure
-grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(60,-3,140,0)),
+grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(56,-2,140,0)),
                          heights = c(1.9,-0.097)),

### 2018-09-12 notebooks/poster_analysis.Rmd
another size test for dendrogram

@@
 # Arrange both plots into one figure
-grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(56,-2,140,0)),
+grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(50,-2,140,0)),
                          heights = c(1.9,-0.097)),

### 2018-09-12 notebooks/poster_analysis.Rmd
adjusted bottom part of dendrogram

@@
 # Arrange both plots into one figure
-grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(50,-2,140,0)),
+grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(50,-2,135,0)),
                          heights = c(1.9,-0.097)),

### 2018-09-12 notebooks/poster_analysis.Rmd
last adjustment to dendrogram

@@
 # Arrange both plots into one figure
-grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(50,-2,135,0)),
+grid.arrange(arrangeGrob(p1 + theme(plot.margin=margin(51,-2,135,0)),
                          heights = c(1.9,-0.097)),

### 2018-09-17 notebooks/poster_analysis.Rmd
added text
added stats for qnr genes

@@
   select(-c(id, type)) %>%
+  mutate(total = rowSums(.),
+         result = if_else(total != 0, 1, 0)) %>%
+  group_by(result) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
+  spread(result, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr))
+```
+
+### Percent qnr genes
+The total percentage of qnr genes identified with ARIBA.
+```{r}
+resfinder_results %>%
+  select_at(.vars = vars("qnrA1","qnrB19",
+                         "qnrB6","qnrB60",
+                         "qnrS1","qnrS2",
+                         "qnrS4")) %>%
   mutate(total = rowSums(.),
@@
 Part of the code below is taken from https://gist.github.com/scwatts/8c32d8f6dc44e352d89ae3de5f3a1125.
+The dendrogram was created by calculating the distance between each allele in the MLST scheme (Escherichia coli #1), and then use the UPGMA method to calculate the dendrogram.
 ```{r fig.height=30, fig.width=10, message=FALSE, warning=FALSE}

### 2018-10-03 src/ariba_functions.R
removed functions file

@@ -1,167 +1,0 @@
-### Collapses data frame to unique lines while ignoring NA's
-func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")
-
-# Confidence interval function
-get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
-                                                c("lwr", "upr")))
-
-# Identifies folder names in input folder
-file_names <- function(filepath) {
-  files <- list.files(path = filepath, pattern = "amr_report")
-  return(files)
-}
-
-# Import ariba data from report.tsv from chosen database used in ariba
-get_ariba_data <- function(filepath) {
-  files <- file_names(filepath)
-  
-  data_list <- lapply(files,
-                      FUN = function(file) {
-                        read.delim(
-                          paste0(filepath, "/", file),
-                          stringsAsFactors = F,
-                          header = TRUE,
-                          sep = "\t"
-                        )
-                      })
-  
-  names(data_list) <- files
-  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
-  return(data)
-}
-
-# Corrects the gene names found in the "cluster" column
-fix_gene_names <- function(df) {
-  genes <- unique(df$ref_name)
-  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
-  new_names <- gsub("_", "", new_names, fixed = T)
-  new_names <- gsub("-", "", new_names, fixed = T)
-  
-  gene_names <- c()
-  for (i in new_names) {
-    p <- paste(tolower(substring(i, 1,3)), substring(i, 4), sep = "", collapse = " ")
-    gene_names <- c(gene_names,p)
-  }
-  df2 <- data.frame(genes,gene_names) %>%
-    mutate(genes = as.character(genes)) %>%
-    rename(ref_name = genes)
-  
-  df <- df %>%
-    left_join(df2, by = "ref_name")
-  
-  return(df)
-}
-
-# Function for selecting genes of interest and filtering the 
-# columns in the data frame on the resulting vector
-select_genes <- function(df, gene_string) {
-  names_df <- names(df)
-  x <- "ref"
-  for (i in names_df) {
-    for (j in gene_string) {
-      if (str_detect(i, regex(j, ignore_case = T)) == TRUE) {
-        x <- c(x, i)
-      }
-    }
-  }
[…]

### 2018-10-03 figure_script.R
created script for making figures for the article

@@ -1,0 +1,128 @@
+## Script for generating figures for the article
+
+gene_names <- names(mut_report)
+
+gene_names <- gene_names[-1]
+
+ggsave(
+  "C:/Users/VI1511/OneDrive - Veterinærinstituttet/Artikler/qrec_wgs/figures/percent_mut_per_species.tiff",
+  mut_report %>%
+    mutate(
+      ref = gsub("(.*?)\\_.+", "\\1", ref),
+      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
+      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
+    ) %>%
+    arrange(desc(ref)) %>%
+    mutate(id = id) %>%
+    left_join(isolate_data, by = "id") %>%
+    gather(key, value, gene_names) %>%
+    group_by(key, value, species) %>%
+    count() %>%
+    ungroup() %>%
+    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+    spread(value, n, fill = 0) %>%
+    rowwise() %>%
+    mutate(
+      Total = Present + Absent,
+      Percent = round(Present / Total * 100, 1),
+      lwr = round(get_binCI(Present, Total)[1], 1),
+      upr = round(get_binCI(Present, Total)[2], 1)
+    ) %>%
+    ggplot(aes(key, Percent, fill = species)) +
+    geom_col(color = "black",
+             position = position_dodge(1)) +
+    geom_errorbar(
+      aes(ymin = lwr, ymax = upr),
+      width = 0.6,
+      position = position_dodge(1)
+    ) +
+    scale_x_discrete(limits = c(
+      "gyrA", "gyrB", "parC", "parE",
+      "marR", "soxS", "tolC"
+    )) +
+    scale_fill_brewer(palette = "Paired") +
+    labs(y = "Percent (%) of isolates") +
+    guides(fill = FALSE) +
+    theme_classic() +
+    theme(
+      axis.text.x = element_text(
+        angle = 90,
+        size = 12,
+        hjust = 1,
+        vjust = 0.4
+      ),
+      axis.title.x = element_blank(),
+      axis.text.y = element_text(size = 12),
+      axis.title.y = element_text(size = 14),
+      strip.text = element_text(size = 12)
+    ) +
+    facet_wrap(~ species),
+  device = "tiff",
+  units = "cm",
[…]

### 2018-10-03 notebooks/article_notebook.Rmd
added rmd for notebook

---
title: "Article data analysis"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable)
```

# Import data
```{r Import data}
isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

no_of_reads <- read.table("../data/read_data.txt",
                          sep = "\t",
                          header = T,
                          stringsAsFactors = F)

```

# ARIBA results analysis
```{r ARIBA analysis}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","aac","oqx")
mut_genes <- c("gyr","par","marR","soxS","tolC","acrR")

# adjust parameters for filtering
if (length(ac_genes) == 1) {
  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
    ac_genes <- "ALL"
    } else {
      ac_genes <- ac_genes
    }
}

if (length(mut_genes) == 1) {  
  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
    mut_genes <- "ALL"
    } else {
      mut_genes <- mut_genes
    }
}
# -------------------------------------- Libraries

if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
               forcats,purrr,stringr,kableExtra,
               knitr,IRdisplay,reprex,svglite)

# -------------------------------------- Functions
# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

# Confidence interval function
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Identifies filenames in input folder
file_names <- function(filepath) {
  files <- list.files(path = filepath, pattern = "amr_report.tsv")
  return(files)
}

# Import ariba data from report.tsv from chosen database used in ariba
get_ariba_data <- function(filepath) {
  files <- file_names(filepath)

  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = TRUE,
                          sep = "\t"
                        )
                      })

  names(data_list) <- files
  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
  return(data)
}

# Corrects the gene names found in the "cluster" column
fix_gene_names <- function(df) {
  genes <- unique(df$ref_name)
  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
  new_names <- gsub("_", "", new_names, fixed = T)
  new_names <- gsub("-", "", new_names, fixed = T)

  gene_names <- c()
  for (i in new_names) {
    p <- paste(tolower(substring(i, 1,3)), substring(i, 4), sep = "", collapse = " ")
    gene_names <- c(gene_names,p)
  }
  df2 <- data.frame(genes,gene_names) %>%
    mutate(genes = as.character(genes)) %>%
    rename(ref_name = genes)

  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))

  return(df)
}

# Function for selecting genes of interest and filtering the 
# columns in the data frame on the resulting vector
select_genes <- function(df, gene_string) {
  names_df <- names(df)
  x <- "ref"
  for (i in names_df) {
    for (j in gene_string) {
      if (str_detect(i, regex(j, ignore_case = T)) == TRUE) {
        x <- c(x, i)
      }
    }
  }
  df <- df %>%
    select(one_of(x))
  return(df)
}

# Allowed flags from the ARIBA report
flag_selection <- c("19","27","147","155","403",
                    "411","915","923","787","795",
                    "531","539","659","667","787","795")

# Function that returns info on flag selection
check_flags <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(flag_result = flag %in% flag_selection,
           flag_result = as.integer(flag_result),
           ref = gsub("(.*?)\\_.+", "\\1", ref),
           ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
           ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)) %>%
    rename("gene" = gene_names)
  return(df)
}

# Function that handles megares data and returns a data frame with information on whether a gene
# is mutated or not. Includes control for QRDR in gyrA.
create_mut_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    filter(flag %in% flag_selection) %>%
    mutate(id = 1:n()) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, flag)) %>%
    gather(gene, mut, -ref) %>%
    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
           result = ifelse(mut != 0, 1, 0),
           result = as.integer(result),
           type = "mut")
  return(df)
}

# Corrects or mutations in QRDR for gyrA, gyrB, parC and parE genes.
# The function returns columns of 1/0 values for whether or not the 
# mutations reported by ARIBA is within the QRDR in the gene:
# gyrA: AA 67 - 106
# gyrB: AA 333 - 481
# ParC: AA 51 - 170
# parE: AA 366 - 523
# Source: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
fix_gyr_par_results <- function(df) {
  df <- df %>%
    mutate(gyrA_result = mut %>% # control for mutation within QRDR for gyrA
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L)), # returns TRUE/FALSE whether value is within range or not
           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result), # filters out results for all other genes
           gyrA_result = as.integer(gyrA_result),
           gyrB_result = mut %>% # control for mutation within QRDR for gyrB
             str_extract_all("\\d+") %>%
             map(as.integer) %>%
             map_lgl(~ any(.x >= 333L & .x <= 481L)),
           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
           gyrB_result = as.integer(gyrB_result),
           parC_result = mut %>% # control for mutation within QRDR for parC
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L)),
           parC_result = if_else(gene != "parC", NA, parC_result),
           parC_result = as.integer(parC_result),
           parE_result = mut %>% # control for mutation within QRDR for parE
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 366L & .x <= 523L)),
           parE_result = if_else(gene != "parE", NA, parE_result),
           parE_result = as.integer(parE_result)) %>%
    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
                                      gene == "gyrB" ~ gyrB_result,
                                      gene == "parC" ~ parC_result,
                                      gene == "parE" ~ parE_result)) %>%
    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"), result_gyr_par, result)) %>%
    select(-c(gyrA_result,
              gyrB_result,
              parC_result,
              parE_result,
              result_gyr_par,
              result))
  return(df)
}

# Function that handles resfinder data and returns a data frame with 
# presence/absence for acquired genes                           
create_acquired_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(id = 1:n()) %>%
    filter(flag %in% flag_selection) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
    select(-c(id, flag)) %>%
    gather(gene, result,-ref) %>%
    mutate(result_total = ifelse(result == "", 0, 1),
           result_total = as.character(result_total),
           type = "gene") %>%
    select(-result)
  return(df)
}

# Function that returns a filtered dataframe based on the string "genes"
filter_mut_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in mut_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = as.character(gene))
  return(df)
}

# Function that returns a filtered dataframe based on the string "acquired_genes"
filter_acquired_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in ac_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene),
           gene = as.character(gene))
  return(df)
}

# calculates percentage of present/absent mutations and genes
calc_stats <- function(df) {
  df <- df %>%
    group_by(gene, result_total) %>%
    count() %>%
    ungroup() %>%
    mutate(result_total = if_else(result_total == 1, "Present", "Absent")) %>%
    spread(result_total, n, fill = 0) %>%
    rowwise() %>%
    mutate(Total = Present + Absent,
           Percent = round(Present/Total*100, 1),
           lwr = round(get_binCI(Present, Total)[1], 1),
           upr = round(get_binCI(Present, Total)[2], 1))
  return(df)
}

# calculates how many mutations are present in the genes
# gyrA, gyrB, parC and parE
calc_no_of_mut <- function(df) {
  df1 <- df %>%
    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
    mutate(entries = sapply(strsplit(.$mut, ","), FUN = function(x) {length(x)})) %>%
    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
    select(-entries) %>%
    gather(id, mut, -c(ref, gene, result, type))

  df2 <- fix_gyr_par_results(df1)

  df3 <- df2 %>%
    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
    filter(test == 1) %>%
    spread(gene, mut) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "0", "", x)) %>%
    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)}),
           mut_gyrB = sapply(strsplit(.$gyrB, ","), FUN = function(x) {length(x)}),
           mut_parC = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)}),
           mut_parE = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
    select(-c(type, id, result_total, test)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "", "0", x))
  return(df3)
}

# creates a data frame with one row per sample, and 1/0 results 
# for mutations in respective genes in columns
create_mut_report <- function(df) {
  df <- df %>%
    select(-mut) %>%
    group_by(ref) %>%
    mutate(id = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, type))
  return(df)
}

# creates a data frame with one row per sample, and 1/0 results
# for acquired genes in columns
create_acquired_report <- function(df) {
  df <- df %>%
    group_by(ref) %>%
    mutate(id = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, type))
  return(df)
}

# prepares data frame for number of mutations heatmap
create_heatmap_df <- function(df) {
  qnr <- names(df)

  qnr_cols <- c()

  for (i in qnr) {
    if (grepl("qnr", i, ignore.case = TRUE) == TRUE) {
      qnr_cols <- c(qnr_cols, i)
    }
  }

  qnr_df <- df %>%
    select_at(vars(ref, qnr_cols)) %>%
    gather(gene, result, -ref) %>%
    mutate(type = "Gene")

  qnr_mut_quant <- mut_quant %>%
    select(-c("gyrA","gyrB","parC","parE")) %>%
    mutate(type = "Mut") %>%
    gather(gene, result, -c(ref, type)) %>%
    mutate(gene = case_when(gene == "mut_gyrA" ~ "gyrA",
                            gene == "mut_gyrB" ~ "gyrB",
                            gene == "mut_parC" ~ "parC",
                            gene == "mut_parE" ~ "parE",
                            TRUE ~ gene)) %>%
    rbind(., qnr_df)
  return(qnr_mut_quant)
}

# -------------------------------------- Analysis
# Import data
mut_data <- get_ariba_data(megares_report_loc)
acquired_data <- get_ariba_data(resfinder_report_loc)

# Clean data
clean_mut_data <- fix_gene_names(mut_data)
clean_acquired_data <- fix_gene_names(acquired_data)

# Check flags
mut_flags <- check_flags(clean_mut_data)
acquired_flags <- check_flags(clean_acquired_data)

# Wrangle
mut_table <- create_mut_table(clean_mut_data)
mut_table_fixed <- fix_gyr_par_results(mut_table)
mut_filtered <- filter_mut_table(mut_table_fixed)
acquired_table <- create_acquired_table(clean_acquired_data)
acquired_filtered <- filter_acquired_table(acquired_table)

# Stats
if (length(mut_genes) == 1) {
  if (mut_genes == "ALL") {
    mut_stats <- calc_stats(mut_table_fixed)
    mut_report <- create_mut_report(mut_table_fixed)
    }
} else {
  mut_stats <- calc_stats(mut_filtered)
  mut_report <- create_mut_report(mut_filtered)
}

if (length(ac_genes) == 1) {
  if (ac_genes == "ALL") {
    acquired_stats <- calc_stats(acquired_table)
    acquired_report <- create_acquired_report(acquired_table)
  }
} else {
  acquired_stats <- calc_stats(acquired_filtered)
  acquired_report <- create_acquired_report(acquired_filtered)
}

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table))
heatmap_df <- create_heatmap_df(acquired_report)

# ---------------------------------- Tables

# number of mutations table
mut_no <- mut_quant %>%
  group_by(mut_gyrA, mut_gyrB, mut_parC, mut_parE) %>%
  count()

# number of isolates per mutation combination
mut_comb <- mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  count()
```

# Tables
## Isolate data
```{r}
isolate_data_complete <- isolate_data %>%
  left_join(no_of_reads, by = "id")

isolate_data_complete
```

## List of genes of interest
```{r}

acquired_ref_names <- acquired_data$ref_name
acquired_ref <- c()

for (i in ac_genes) {
  for (j in acquired_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      acquired_ref <- c(acquired_ref, j)
    }
  }
}

acquired_ref <- unique(acquired_ref)

acquired_ref_genes <- acquired_data %>%
  mutate(test = duplicated(ref_name),
         gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
                          grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
                          grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
                          grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
                          grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
                          grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
                          grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9")) %>%
  filter(test == FALSE,
         ref_name %in% acquired_ref_names) %>%
  select(ref_name, gene) %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
         Type = "Acquired")

df_ref_names <- mut_data$cluster
ref_names <- c()

for (i in mut_genes) {
  for (j in df_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      ref_names <- c(ref_names, j)
    }
  }
}

ref_names <- unique(ref_names)

intrinsic_ref_genes <- mut_data %>%
  mutate(test = duplicated(cluster),
         gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
                          grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
                          grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
                          grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
                          grepl("soxs", cluster, ignore.case = T) == TRUE ~ "soxS",
                          grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
                          grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR")) %>%
  filter(test == FALSE,
         cluster %in% ref_names) %>%
  select(ref_name, gene, free_text) %>%
  mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
  select(-ref_name) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
                                 Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "gyrB" ~ "DNA gyrase subunit B, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
                                 Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
                                 Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
                                 Gene == "soxS" ~ "Transcriptional activator of the superoxide response regulon.",
                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump."),
         Type = "Intrinsic gene")

ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)

ref_genes

```

```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  count()

per_species
```

### 2018-10-03 ariba_script.R
integrated into notebook

@@ -1,21 +1,353 @@
-# Libraries
+#!/usr/bin/env Rscript
+args <- commandArgs(trailingOnly = TRUE)
+
+megares_report_loc <- "data/ariba/megares_results"
+resfinder_report_loc <- "data/ariba/resfinder_results"
+ac_genes <- c("qnr","aac","oqx")
+mut_genes <- c("gyr","par","marR","soxS","tolC","acrR")
+
+# adjust parameters for filtering
+if (length(ac_genes) == 1) {
+  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
+    ac_genes <- "ALL"
+    } else {
+      ac_genes <- ac_genes
+    }
+}
+
+if (length(mut_genes) == 1) {  
+  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
+    mut_genes <- "ALL"
+    } else {
+      mut_genes <- mut_genes
+    }
+}
+# -------------------------------------- Libraries
+
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,cowplot,forcats,R.utils,purrr,stringr,kableExtra,knitr,IRdisplay,reprex)
+pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
+               forcats,purrr,stringr,kableExtra,
+               knitr,IRdisplay,reprex,svglite)
 
-# Import functions
-source("src/ariba_functions.R")
+# -------------------------------------- Functions
+# Collapses data frame to unique lines while ignoring NA's
+func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")
 
-# Set input data path
-megares_results <- "data/ariba/megares_results"
-resfinder_results <- "data/ariba/resfinder_results"
+# Confidence interval function
+get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
+                                                c("lwr", "upr")))
 
-# Specify genes of interest
-genes <- c("gyr", "par", "sox", "mar", "acr", "tol", "omp")
-acquired_genes <- c("qnr", "bla", "tet", "sul", "oqx")
+# Identifies filenames in input folder
+file_names <- function(filepath) {
+  files <- list.files(path = filepath, pattern = "amr_report.tsv")
+  return(files)
+}
 
+# Import ariba data from report.tsv from chosen database used in ariba
+get_ariba_data <- function(filepath) {
+  files <- file_names(filepath)
+  
[…]

### 2018-10-03 notebooks/poster_analysis.Rmd
minor changes

@@
 # Description
-This notebook presents the analysis of whole genome data for the poster "Quinolone resistance mechanisms found in E. coli from four animal species in Norway
-", presented at the ASM NGS Conference in Washington D.C. September 2018.
+This notebook presents the analysis of whole genome data for the poster "Quinolone resistance mechanisms found in E.
+coli from four animal species in Norway", presented at the ASM NGS Conference in Washington D.C.
+September 2018.
 
@@
                            header = T,
-                           stringsAsFactors = F) %>%
+                           stringsAsFactors = F)
   filter(id %not_in% c("2016-02-428-2-S","2016-02-486-2-S","2016-02-732-2-S","2016-17-565-1-S"))
@@
                            header = T,
-                           stringsAsFactors = F) %>%
-  filter(id %not_in% c("2016-02-428-2-S","2016-02-486-2-S","2016-02-732-2-S","2016-17-565-1-S"))
+                           stringsAsFactors = F)
 
@@
              ncol=2)
-
 ```

### 2018-10-04 notebooks/article_notebook.Rmd
minor changes

@@
 
+# Dependencies
+This notebook is using the following packages:
+
+- dplyr
+
+- ggplot2
+
+- purrr
 ```{r load libraries, message=FALSE, warning=FALSE}
@@
 
-# Import data
-```{r Import data}
+```{r, include=FALSE}
+id <- c("2017-01-3941-1-S", "2016-17-98-1-S", "2016-17-610-1-S", "2016-17-604-1-S", 
+"2016-17-565-1-S", "2016-17-550-1-S", "2016-17-548-1-S", "2016-17-546-1-S", 
+"2016-17-485-1-S", "2016-17-469-1-S", "2016-17-423-1-S", "2016-17-412-1-S", 
+"2016-17-408-1-T", "2016-17-408-1-S", "2016-17-393-1-S", "2016-17-391-1-S", 
+"2016-17-388-1-T", "2016-17-388-1-S", "2016-17-377-1-S", "2016-17-363-1-S", 
+"2016-17-352-1-S", "2016-17-343-1-S", "2016-17-340-1-S", "2016-17-336-1-S", 
+"2016-17-329-1-S", "2016-17-299-1-S", "2016-17-298-1-S", "2016-17-294-1-S", 
+"2016-17-292-1-S", "2016-17-288-1-S", "2016-17-281-1-T", "2016-17-281-1-S", 
+"2016-17-275-1-S", "2016-17-271-1-S", "2016-17-27-1-S", "2016-17-264-1-S", 
+"2016-17-263-1-S", "2016-17-257-1-S", "2016-17-256-1-S", "2016-17-242-1-S", 
+"2016-17-225-1-S", "2016-17-224-1-S", "2016-17-189-1-T", "2016-17-183-1-S", 
+"2016-17-174-1-T", "2016-17-174-1-S", "2016-17-164-1-S", "2016-17-16-1-S", 
+"2016-17-148-1-S", "2016-17-124-1-S", "2016-17-114-1-S", "2016-17-105-1-T", 
+"2016-17-1-1-S", "2016-02-785-2-S", "2016-02-783-2-T", "2016-02-771-2-S", 
+"2016-02-770-2-S", "2016-02-741-2-S", "2016-02-739-2-S", "2016-02-732-2-S", 
+"2016-02-727-2-S", "2016-02-726-2-S", "2016-02-725-2-S", "2016-02-723-2-S", 
+"2016-02-697-2-S", "2016-02-693-2-T", "2016-02-693-2-S", "2016-02-692-2-S", 
+"2016-02-690-2-T", "2016-02-690-2-S", "2016-02-620-2-S", "2016-02-563-2-S", 
+"2016-02-546-2-S", "2016-02-543-2-S", "2016-02-531-2-S", "2016-02-522-2-S", 
+"2016-02-515-2-T", "2016-02-515-2-S", "2016-02-514-2-S", "2016-02-513-2-T", 
+"2016-02-513-2-S", "2016-02-505-2-S", "2016-02-492-2-S", "2016-02-490-2-S", 
+"2016-02-486-2-S", "2016-02-447-2-S", "2016-02-446-2-S", "2016-02-428-2-S", 
+"2016-02-425-2-S", "2016-02-419-2-S", "2016-02-418-2-S", "2016-02-415-2-S", 
+"2016-02-413-2-T", "2016-02-413-2-S", "2016-02-411-2-S", "2016-02-410-2-S", 
+"2016-02-409-2-S", "2016-02-408-2-S", "2016-02-402-2-S", "2016-02-381-2-S", 
+"2016-02-369-2-S", "2016-02-366-2-S", "2016-02-363-2-S", "2016-02-359-2-S", 
+"2016-02-358-2-S", "2016-02-356-2-S", "2016-02-353-2-S", "2016-02-352-2-S", 
+"2016-02-351-2-S", "2016-02-348-2-S", "2016-02-345-2-T", "2016-02-345-2-S", 
+"2016-02-343-2-S", "2016-02-336-2-S", "2016-02-330-2-S", "2016-02-329-2-S", 
+"2016-02-324-2-S", "2016-02-316-2-S", "2016-02-302-2-S", "2016-02-293-2-S", 
+"2016-02-281-2-S", "2016-02-278-2-S", "2016-01-3763-1-T", "2016-01-3474-1-T", 
+"2015-01-753-1-S", "2015-01-752-1-S", "2015-01-751-1-S", "2015-01-750-1-S", 
+"2015-01-749-1-S", "2015-01-5263-1-S", "2015-01-5262-1-S", "2015-01-5177-1-S", 
+"2015-01-5141-1-S", "2015-01-5140-1-S", "2015-01-5138-1-S", "2015-01-5022-1-S", 
+"2015-01-5020-1-S", "2015-01-5018-1-S", "2015-01-4988-1-S", "2015-01-4934-1-S", 
+"2015-01-4932-1-S", "2015-01-4931-1-S", "2015-01-4794-1-S", "2015-01-4788-1-S", 
+"2015-01-4742-1-S", "2015-01-467-1-S", "2015-01-466-1-S", "2015-01-463-1-S", 
+"2015-01-4559-1-S", "2015-01-4558-1-S", "2015-01-4465-1-S", "2015-01-4463-1-S", 
+"2015-01-4462-1-S", "2015-01-4421-1-S", "2015-01-4249-1-S", "2015-01-4037-1-S", 
+"2015-01-3915-1-S", "2015-01-3914-1-S", "2015-01-3911-1-S", "2015-01-3887-1-S", 
+"2015-01-3640-1-S", "2015-01-3639-1-S", "2015-01-3638-1-S", "2015-01-3548-1-S", 
+"2015-01-3545-1-S", "2015-01-3318-1-S", "2015-01-3317-1-S", "2015-01-3102-1-S", 
+"2015-01-3004-1-S", "2015-01-3001-1-S", "2015-01-2879-1-S", "2015-01-2878-1-S", 
+"2015-01-2839-1-S", "2015-01-2838-1-S", "2015-01-2837-1-S", "2015-01-2836-1-S", 
+"2015-01-2835-1-S", "2015-01-2771-1-S", "2015-01-2430-1-S", "2015-01-2429-1-S", 
+"2015-01-2426-1-S", "2015-01-2404-1-S", "2015-01-2403-1-S", "2015-01-2309-1-S", 
+"2015-01-2166-1-S", "2015-01-2097-1-S", "2015-01-2094-1-S", "2015-01-1955-1-S", 
+"2015-01-1954-1-S", "2015-01-1952-1-T", "2015-01-1951-1-S", "2015-01-1950-1-S", 
+"2015-01-1949-1-S", "2015-01-1826-1-S", "2015-01-1823-1-S", "2015-01-1661-1-S", 
+"2015-01-1566-1-S", "2015-01-1343-1-S", "2014-01-861-1-S", "2014-01-860-1-S", 
+"2014-01-856-1-S", "2014-01-7584-1-S", "2014-01-7543-1-S", "2014-01-7512-1-S", 
+"2014-01-7490-1-S", "2014-01-7479-1-S", "2014-01-7443-1-S", "2014-01-7442-1-S", 
+"2014-01-7394-1-S", "2014-01-7375-1-S", "2014-01-7234-1-S", "2014-01-7207-1-S", 
+"2014-01-7136-1-S", "2014-01-7133-1-S", "2014-01-7050-1-S", "2014-01-6926-1-S", 
+"2014-01-6924-1-S", "2014-01-6710-1-S", "2014-01-6625-1-S", "2014-01-6530-1-S", 
+"2014-01-6291-1-S", "2014-01-6123-1-S", "2014-01-6122-1-S", "2014-01-6044-1-S", 
+"2014-01-6043-1-S", "2014-01-6041-1-S", "2014-01-6040-1-S", "2014-01-5917-1-T", 
+"2014-01-5917-1-S", "2014-01-5916-1-S", "2014-01-5915-1-S", "2014-01-5912-1-S", 
+"2014-01-5792-1-S", "2014-01-5565-1-S", "2014-01-4922-1-T", "2014-01-4921-1-S", 
+"2014-01-4651-1-S", "2014-01-4539-1-S", "2014-01-4271-1-S", "2014-01-4270-1-S", 
+"2014-01-3802-1-S", "2014-01-3677-1-S", "2014-01-3676-1-S", "2014-01-3212-1-S", 
+"2014-01-2996-1-S", "2014-01-2995-1-S", "2014-01-2992-1-S", "2014-01-2773-1-S", 
+"2014-01-2728-1-S", "2014-01-2516-1-S", "2014-01-2366-1-S", "2014-01-2362-1-T", 
+"2014-01-2145-1-S", "2014-01-2144-1-S", "2014-01-2071-1-S", "2014-01-2070-1-S", 
+"2014-01-2069-1-S", "2014-01-1742-1-S", "2014-01-1741-1-S", "2014-01-1649-1-S", 
+"2014-01-1268-1-T", "2014-01-1182-1-S", "2013-01-2024-1-T", "2012-01-838-6-T", 
+"2011-01-9386-3-T", "2011-01-6927-5-T", "2011-01-3460-5-T", "2011-01-1544-2-T", 
+"2010-01-991-7-T", "2009-01-551-1-T", "2009-01-4672-5-T", "2009-01-4618-2-T", 
+"2009-01-4415-4-T", "2009-01-413-1-T", "2009-01-4064-1-T", "2009-01-3815-1-T", 
+"2009-01-3064-2-T", "2009-01-2696-4-T", "2009-01-2521-3-T", "2009-01-175-4-T", 
+"2009-01-175-2-T", "2009-01-1690-1-T", "2007-01-361-1-T", "2006-01-2821-1-T", 
+"2006-01-2184-1-T")
+```
+
+```{r Import data, include = FALSE}
 isolate_data <- read.table("../data/isolate_data.txt",
@@
 
-# ARIBA results analysis
-```{r ARIBA analysis}
+```{r ARIBA analysis, include = FALSE}
 megares_report_loc <- "../data/ariba/megares_results"
@@
 
+## Isolates per species
 ```{r}
@@
 ```
+
+# Figures
+## Percent QREC isolates with mutations in intrinsic genes
+```{r}
+gene_names <- names(mut_report)
+
+gene_names <- gene_names[-1]
+
+mut_report %>%
+    mutate(
+      ref = gsub("(.*?)\\_.+", "\\1", ref),
+      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
+      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
+    ) %>%
+    arrange(desc(ref)) %>%
+    mutate(id = id) %>%
+    left_join(isolate_data, by = "id") %>%
+    gather(key, value, gene_names) %>%
+    group_by(key, value, species) %>%
+    count() %>%
+    ungroup() %>%
+    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+    spread(value, n, fill = 0) %>%
+    rowwise() %>%
+    mutate(
+      Total = Present + Absent,
+      Percent = round(Present / Total * 100, 1),
+      lwr = round(get_binCI(Present, Total)[1], 1),
+      upr = round(get_binCI(Present, Total)[2], 1)
+    ) %>%
+    ggplot(aes(key, Percent, fill = species)) +
+    geom_col(color = "black",
+             position = position_dodge(1)) +
+    geom_errorbar(
+      aes(ymin = lwr, ymax = upr),
+      width = 0.6,
+      position = position_dodge(1)
+    ) +
+    scale_x_discrete(limits = c(
+      "gyrA", "gyrB", "parC", "parE",
+      "marR", "soxS", "tolC"
+    )) +
+    scale_fill_brewer(palette = "Paired") +
+    labs(y = "Percent (%) of isolates") +
+    guides(fill = FALSE) +
+    theme_classic() +
+    theme(
+      axis.text.x = element_text(
+        angle = 90,
+        size = 12,
+        hjust = 1,
+        vjust = 0.4
+      ),
+      axis.title.x = element_blank(),
+      axis.text.y = element_text(size = 12),
+      axis.title.y = element_text(size = 14),
+      strip.text = element_text(size = 12)
+    ) +
+    facet_wrap(~ species)
+```
+
+## Percent QREC isolates with Qnr genes per species
+```{r}
+qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")
+
+acquired_report %>%
+    mutate(
+      ref = gsub("(.*?)\\_.+", "\\1", ref),
+      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
+      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
+    ) %>%
+    arrange(desc(ref)) %>%
+    mutate(id = id) %>%
+    left_join(isolate_data, by = "id") %>%
+    gather(key, value, qnr_genes) %>%
+    group_by(key, value, species) %>%
+    count() %>%
+    ungroup() %>%
+    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
+    spread(value, n, fill = 0) %>%
+    rowwise() %>%
+    mutate(
+      Total = Present + Absent,
+      Percent = round(Present / Total * 100, 1),
+      lwr = round(get_binCI(Present, Total)[1], 1),
+      upr = round(get_binCI(Present, Total)[2], 1)
+    ) %>%
+    ggplot(aes(key, Percent, fill = species)) +
+    geom_col(color = "black",
+             position = position_dodge(0.9)) +
+    geom_errorbar(
+      aes(ymin = lwr, ymax = upr),
+      width = 0.6,
+      position = position_dodge(0.9)
+    ) +
+    scale_fill_brewer(palette = "Paired") +
+    labs(y = "Percent (%) of isolates",
+         fill = NULL) +
+    theme_classic() +
+    theme(
+      axis.text = element_text(size = 12),
+      axis.title.y = element_text(size = 14),
+      legend.text = element_text(size = 12),
+      axis.title.x = element_blank(),
+      legend.justification = c(0, 1),
+      legend.position = c(0.82, 0.97)
+    )
+```

### 2018-10-04 notebooks/fastqc_analysis.Rmd
initial commit

---
title: "Fastqc Analysis of Raw Reads"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

# Dependencies
This analysis depends on the following packages:

- fastqcr

- dplyr

- ggplot2

- tidyr

- viridis

- ggsci

- scales

```{r, include = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(fastqcr,dplyr,ggplot2,tidyr,viridis,ggsci,scales)
```

```{r, include = FALSE}
input_dir <- "C:/Users/VI1511/OneDrive - Veterinærinstituttet/R/R_Projects/qrec_wgs1/data/fastqc"
```

# Functions
```{r}
# Function that lists all the file names in the input folder
file_names <- function(filepath, folder) {
  files <- list.files(path = paste0(filepath, "/", folder), pattern = "_fastqc.zip")
  return(files)
}

# Function that searches recursively for all filenames with fastqc.zip
file_names_recursive <- function(filepath) {
  files <- list.files(path = filepath, pattern = "_fastqc.zip", recursive = TRUE)
  return(files)
}

# Lists folders in the input folder
folder_names <- function(filepath) {
  folders <- list.files(path = filepath)
  return(folders)
}

# Identifies the names of the files and groups them accoring to their respective folders
get_grouping_variable <- function(path, folder) {
  files <- file_names(path, folder)
  df <- data.frame(files = files, group = folder, stringsAsFactors = FALSE)
  df$files <- gsub("(.*?)_fastqc.zip", "\\1", df$files)
  colnames(df) <- c("ref","group")
  return(df)
}

# Creates a data frame with grouping information of the reports
create_group_df <- function(path) {
  folders <- folder_names(path)
  if (length(folders) > 1) {
    df <- lapply(folders, function(folder) get_grouping_variable(path, folder))
    df <- bind_rows(df)
  } else {
    df <- get_grouping_variable(path, folders)
  }
  return(df)
}

# Function that filter out duplicated counts
filter_counts <- function(df) {
  df <- df %>%
    mutate(dupl = duplicated(Count)) %>%
    filter(dupl == FALSE)
  return(df)
}

# Function that prepares the sequence length data for plotting
prepare_seq_len_data <- function(list) {
  x <- split(list$sequence_length_distribution, list$sequence_length_distribution$group)
  x <- lapply(x, filter_counts)
  x <- bind_rows(x)
  return(x)
}

# Function that imports and wrangles fastqc data
get_fastqc_data <- function(filepath) {
  folders <- folder_names(filepath)

  get_files <- file_names_recursive(filepath)

  data_list <- lapply(get_files,
                      FUN = function(file) {
                        qc_read(paste0(filepath, "/", file),
                                modules = "all",
                                verbose = FALSE)
                      })
  names(data_list) <- gsub("(.*?)/(.*?)_fastqc.zip", "\\2", get_files)
  data_list <- purrr::transpose(data_list)

  data_list$sequence_length_distribution <- NULL
  data_list$kmer_content <- NULL

  list_names <- names(data_list)
  list_numbers <- 1:length(list_names)
  for (i in list_numbers) {
    assign(list_names[i], bind_rows(data_list[[i]], .id = "ref"))
  }
  df_list <- list(summary,
                  basic_statistics,
                  per_base_sequence_quality,
                  per_tile_sequence_quality,
                  per_sequence_quality_scores,
                  per_base_sequence_content,
                  per_sequence_gc_content,
                  per_base_n_content,
                  sequence_duplication_levels,
                  overrepresented_sequences,
                  adapter_content,
                  total_deduplicated_percentage)
  names(df_list) <- list_names
  df_list$basic_statistics <- df_list$basic_statistics %>%
    spread(Measure,Value) %>%
    left_join(group_df, by = "ref")
  return(df_list)
}
```

# Grouping
This function creates a data frame with the grouping information of the files, based on the folder names that hold them. Each file is specified a group name based on the folder they are in. This information is used to specify each group of files in the plots below.

```{r}
group_df <- create_group_df(input_dir)
```

# Import data
This function imports the data from the location specified.
```{r}
df_list <- get_fastqc_data(input_dir)
```

# Plotting

## Adapter content
This plot presents the adapter content for all sequences in each group. Each adapter type is specified by color. The X-axis represents the read positions, where the left side represents the start of the read and the right side the end.

```{r}
df_list$adapter_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    # get data ready for boxplot by gathering the columns into two columns
    gather(key,
           value, -c(ref,
                     Position,
                     group,
                     seqlen)) %>%
    # plotting
    ggplot(aes(factor(
      Position,
      levels = unique(Position),
      ordered = TRUE
    ), value, color = key)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(
      x = "Position in Read",
      y = "Percent (%) Adapter Content",
      color = NULL,
      title = "Adapter content"
    ) +
    scale_colour_jama() +
    scale_y_continuous(limits = c(0, 20)) +
    theme_classic() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "bottom"
    ) +
    # creates separate windows for each group
    facet_wrap(~ group, scales = "free", dir = "v")
```

## Per base sequence content
This plot presents the per base sequence content for all reads in each group. Each base is represented by color. 
```{r}
df_list$per_base_sequence_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    gather(key, value, -c(ref, Base, group, seqlen)) %>%
    ggplot(aes(factor(
      Base, levels = unique(Base), ordered = TRUE
    ), value, color = key)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(
      x = "Position in Read",
      y = "Percent (%)",
      color = NULL,
      title = "Per base sequence content"
    ) +
    theme_classic() +
    scale_color_jama() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "bottom"
    ) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free", dir = "v")
```

## Sequence duplication levels
This plot presents the duplication levels for all reads in each group. The X-axis contain "bins" of different duplication levels, and the number is a representation of how many of the same read is present. For example, the bin "3" represents the reads that have triplicates in the file.

```{r}
df_list$sequence_duplication_levels %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    gather(key, value, -c(ref, `Duplication Level`, group, seqlen)) %>%
    ggplot(aes(
      factor(
        `Duplication Level`,
        levels = unique(`Duplication Level`),
        ordered = TRUE
      ),
      value,
      fill = key
    )) +
    geom_boxplot(outlier.size = 0.5) +
    scale_fill_manual(values = c("#ef8a62",
                                 "#67a9cf")) +
    theme_classic() +
    labs(x = "Duplication Level",
         y = "Percent (%) of Sequences",
         fill = NULL,
         title = "Sequence duplication levels") +
    theme(legend.position = "bottom",
          axis.text.x = element_text(
            angle = 90,
            hjust = 1,
            vjust = 0.4
          )) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free")
```

## Sequence quality scores
This plot presents the sequence quality scores for all reads in each group. The X-axis represent the phred-score for each read, and the y-axis represent the amount of reads per phred-score. Higher score is better.

```{r}
df_list$per_sequence_quality_scores %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(Quality), Count, fill = factor(Quality))) +
    geom_boxplot(outlier.size = 0.5) +
    scale_y_continuous(labels = comma) +
    scale_x_discrete(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40)) +
    scale_fill_viridis(discrete = TRUE) +
    labs(x = "Quality",
         y = "Number of reads",
         title = "Per sequence quality scores") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 10)) +
    # creates separate windows for each group
    facet_wrap( ~ group)
```

## Per sequence GC-content
This plot presents the % GC-content for all reads in each group.
```{r}
df_list$per_sequence_gc_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(`GC Content`), Count)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(x = "GC content (%)",
         y = "Number of reads",
         title = "Per sequence GC content") +
    scale_y_continuous(labels = comma) +
    scale_x_discrete(breaks = as.character(seq(
      from = 0, to = 100, by = 10
    ))) +
    theme_classic() +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free")
```

## Per base N-content
This plot presents the per base N content for all reads in each group. N's are added to the read instead of a base if the sequencer is unable to do the base call with sufficient confidence. As before, the X-axis represents the position in the read, where left is the start of the read and right is the end.
```{r}
df_list$per_base_n_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(
      Base, levels = unique(Base), ordered = TRUE
    ), `N-Count`)) +
    geom_boxplot(fill = "#e6e6e6",
                 outlier.size = 0.5) +
    labs(x = "Position in read",
         title = "Per base N content") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free", dir = "v")
```

## Total deduplicated percentage
This plot presents the total deduplicated percentage for all reads in each group.
```{r}
df_list$total_deduplicated_percentage %>%
    gather(key = "ref", value = `1`) %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length",
           perc = `1`) %>%
    mutate(perc = as.numeric(perc)) %>%
    ggplot(aes(factor(group), perc)) +
    geom_boxplot(fill = "#e6e6e6",
                 outlier.size = 0.5) + 
    scale_y_continuous(limits = c(0, 100)) +
    labs(x = "Group",
         y = "Total percentage of deduplicated reads",
         title = "Total deduplicated percentage") +
    theme_classic()
```

## Per base sequence quality
This plot presents the per base mean sequence quality for all reads in each group. The green line depicts the "good quality" limit of the phred-score 28. As before, the X-axis represents the position in each read.
```{r}
df_list$per_base_sequence_quality %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    mutate(Base = factor(Base,
                         levels = unique(Base),
                         ordered = TRUE)) %>%
    group_by(group) %>%
    mutate(xmax = length(unique(Base)) + 1) %>%
    ungroup() %>%
    ggplot(aes(Base,
               Mean)) +
    geom_boxplot(outlier.size = 0.4,
                 fill = "#7f7f7f") +
    geom_hline(aes(yintercept = 28),
               color = "green") +
    labs(x = "Position in read",
         y = "Sequence quality",
         title = "Per base mean sequence quality") +
    scale_y_continuous(limits = c(0, 42)) +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    # creates separate windows for each group
    facet_wrap(~ group, scales = "free", dir = "v")
```

### 2018-10-04 notebooks/fastqc_analysis.Rmd
changed size of plots and x/y axis scales

@@
 
-```{r}
+```{r, fig.height=8, fig.width=12}
 df_list$adapter_content %>%
@@
     # creates separate windows for each group
-    facet_wrap(~ group, scales = "free", dir = "v")
+    facet_wrap(~ group, scales = "free_x", dir = "v")
+
 ```
@@
 This plot presents the per base sequence content for all reads in each group. Each base is represented by color. 
-```{r}
+```{r, fig.height=8, fig.width=12}
 df_list$per_base_sequence_content %>%
@@
     # creates separate windows for each group
-    facet_wrap( ~ group, scales = "free", dir = "v")
+    facet_wrap( ~ group, scales = "free_x", dir = "v")
 ```
@@
 
-```{r}
+```{r, fig.height=8, fig.width=12}
 df_list$sequence_duplication_levels %>%
@@
     # creates separate windows for each group
-    facet_wrap( ~ group, scales = "free")
+    facet_wrap( ~ group, scales = "free_x")
 ```
@@
 
-```{r}
+```{r, fig.height=8, fig.width=12}
 df_list$per_sequence_quality_scores %>%
@@
 This plot presents the % GC-content for all reads in each group.
-```{r}
+```{r, fig.height=8, fig.width=12}
 df_list$per_sequence_gc_content %>%
@@
 As before, the X-axis represents the position in the read, where left is the start of the read and right is the end.
-```{r}
+```{r, fig.height=8, fig.width=12}
 df_list$per_base_n_content %>%
@@
 This plot presents the total deduplicated percentage for all reads in each group.
-```{r}
+```{r, fig.height=8, fig.width=12}
 df_list$total_deduplicated_percentage %>%
@@
 As before, the X-axis represents the position in each read.
-```{r}
+```{r, fig.height=8, fig.width=12}
 df_list$per_base_sequence_quality %>%
@@
     # creates separate windows for each group
-    facet_wrap(~ group, scales = "free", dir = "v")
+    facet_wrap(~ group, scales = "free_x", dir = "v")
 ```

### 2018-10-04 notebooks/fastqc_analysis.Rmd
updated figure text size

@@
   gather(key,
-           value, -c(ref,
+  value,-c(ref,
   Position,
@@
   axis.text.x = element_blank(),
+  axis.title = element_text(size = 14),
   axis.ticks.x = element_blank(),
+  strip.text = element_text(size = 14),
+  axis.text.y = element_text(size = 12),
+  legend.text = element_text(size = 12),
   legend.position = "bottom"
@@
   # creates separate windows for each group
-    facet_wrap(~ group, scales = "free_x", dir = "v")
-
+  facet_wrap( ~ group, scales = "free_x", dir = "v")
 ```
@@
   rename(seqlen = "Sequence length") %>%
-    gather(key, value, -c(ref, Base, group, seqlen)) %>%
+  gather(key, value,-c(ref, Base, group, seqlen)) %>%
   ggplot(aes(factor(
@@
   axis.ticks.x = element_blank(),
+  axis.title = element_text(size = 14),
+  strip.text = element_text(size = 14),
+  axis.text.y = element_text(size = 12),
+  legend.text = element_text(size = 12),
   legend.position = "bottom"
@@
   # creates separate windows for each group
-    facet_wrap( ~ group, scales = "free_x", dir = "v")
+  facet_wrap(~ group, scales = "free_x", dir = "v")
 ```
@@
   rename(seqlen = "Sequence length") %>%
-    gather(key, value, -c(ref, `Duplication Level`, group, seqlen)) %>%
+  gather(key, value,-c(ref, `Duplication Level`, group, seqlen)) %>%
   ggplot(aes(
@@
   title = "Sequence duplication levels") +
-    theme(legend.position = "bottom",
+  theme(
+  axis.ticks.x = element_blank(),
+  axis.title = element_text(size = 14),
+  strip.text = element_text(size = 14),
+  axis.text.y = element_text(size = 12),
+  legend.text = element_text(size = 12),
+  legend.position = "bottom",
   axis.text.x = element_text(
+  size = 12,
   angle = 90,
@@
   vjust = 0.4
-          )) +
+  )
+  ) +
   # creates separate windows for each group
-    facet_wrap( ~ group, scales = "free_x")
+  facet_wrap(~ group, scales = "free_x")
 ```
@@
   theme_classic() +
-    theme(axis.text.x = element_text(size = 10)) +
+  theme(
+  axis.text = element_text(size = 12),
+  axis.title = element_text(size = 14),
+  strip.text = element_text(size = 14)
+  ) +
   # creates separate windows for each group
-    facet_wrap( ~ group)
+  facet_wrap(~ group)
 ```
@@
   theme_classic() +
+  theme(
+  axis.text = element_text(size = 12),
+  axis.title = element_text(size = 14),
+  strip.text = element_text(size = 14)
+  ) +
   # creates separate windows for each group
-    facet_wrap( ~ group, scales = "free_x")
+  facet_wrap(~ group, scales = "free_x")
 ```
@@
   theme_classic() +
-    theme(axis.text.x = element_blank(),
-          axis.ticks.x = element_blank()) +
+  theme(
+  axis.text.x = element_blank(),
+  axis.ticks.x = element_blank(),
+  axis.title = element_text(size = 14),
+  strip.text = element_text(size = 14),
+  axis.text.y = element_text(size = 12)
+  ) +
   # creates separate windows for each group
-    facet_wrap( ~ group, scales = "free_x", dir = "v")
+  facet_wrap(~ group, scales = "free_x", dir = "v")
 ```
@@
   title = "Total deduplicated percentage") +
-    theme_classic()
+  theme_classic() +
+  theme(axis.text = element_text(size = 12),
+  axis.title = element_text(size = 14))
 ```
@@
   theme_classic() +
-    theme(axis.text.x = element_blank(),
-          axis.ticks.x = element_blank()) +
+  theme(
+  axis.text.x = element_blank(),
+  axis.ticks.x = element_blank(),
+  axis.title = element_text(size = 14),
+  strip.text = element_text(size = 14),
+  axis.text.y = element_text(size = 12)
+  ) +
   # creates separate windows for each group
-    facet_wrap(~ group, scales = "free_x", dir = "v")
+  facet_wrap( ~ group, scales = "free_x", dir = "v")
 ```

### 2018-10-08 notebooks/article_notebook.Rmd
updated with a new figure

@@
 
+mic_data <- read.table("../data/mic_data.txt",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
+
 ```
@@
 resfinder_report_loc <- "../data/ariba/resfinder_results"
-ac_genes <- c("qnr","aac","oqx")
-mut_genes <- c("gyr","par","marR","soxS","tolC","acrR")
+ac_genes <- c("qnr","oqx")
+mut_genes <- c("gyr","par","marR","soxR","acrR")
 
@@
 
+# Introduces unique ID's for each sample
+fix_ref_id <- function(df) {
+  df <- df %>%
+    mutate(ref = gsub("(.*?)\\_.+", "\\1", ref),
+         ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
+         ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)) %>%
+    arrange(desc(ref)) %>%
+    mutate(id = id) %>%
+    select(-ref)
+  return(df)
+}
+
 # -------------------------------------- Analysis
@@
     mut_report <- create_mut_report(mut_table_fixed)
+    mut_report_fixed <- fix_ref_id(mut_report)
     }
@@
   mut_report <- create_mut_report(mut_filtered)
+  mut_report_fixed <- fix_ref_id(mut_report)
 }
@@
     acquired_report <- create_acquired_report(acquired_table)
+    acquired_report_fixed <- fix_ref_id(acquired_report)
   }
@@
   acquired_report <- create_acquired_report(acquired_filtered)
+  acquired_report_fixed <- fix_ref_id(acquired_report)
 }
@@
 mut_quant <- suppressWarnings(calc_no_of_mut(mut_table))
+mut_quant_fixed <- fix_ref_id(mut_quant)
 heatmap_df <- create_heatmap_df(acquired_report)
@@
 
+## Percent isolates with PMQR and/or QRDR mutations in gyrA, gyrB, parC or parE
+```{r}
+total_report <- mut_report_fixed %>%
+  left_join(acquired_report_fixed, by = "id") %>%
+  gather(gene, value, -id) %>%
+  mutate(present = ifelse(value == 1, gene, NA),
+         ref = 1:n()) %>%
+  spread(gene, value) %>%
+  group_by(id) %>%
+  summarise_all(funs(func_paste)) %>%
+  select(-ref) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  group_by(present,CIP) %>%
+  count() %>%
+  ungroup() %>%
+  rowwise() %>%
+  mutate(present = ifelse(present == "", "None", present),
+         total = 285,
+         percent = round(n/total * 100, 1),
+         lwr = round(get_binCI(n, total)[1], 1),
+         upr = round(get_binCI(n, total)[2], 1))
+
+genes <- c(names(mut_report_fixed), names(acquired_report_fixed))
+genes <- genes[!genes == "id"]
+
+p1 <- ggplot(total_report, aes(reorder(present, as.numeric(-CIP)), percent, fill = factor(CIP)))+
+  geom_col(color = "black")+
+  labs(y = "Percent (%) isolates")+
+  scale_fill_viridis(discrete = TRUE)+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_text(size = 10),
+        axis.text.y = element_text(size = 8),
+        axis.title.x = element_blank(),
+        legend.text = element_text(size = 8),
+        legend.title = element_blank(),
+        legend.position = "top",
+        legend.spacing.x = unit(0.27, "cm"))
+
+dot_report <- mut_report_fixed %>%
+  left_join(acquired_report_fixed, by = "id") %>%
+  gather(gene, value, -id) %>%
+  mutate(present = ifelse(value == 1, gene, NA),
+         ref = 1:n()) %>%
+  spread(gene, value) %>%
+  group_by(id) %>%
+  summarise_all(funs(func_paste)) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  select(-c(ref, id)) %>%
+  ungroup() %>%
+  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
+  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
+         present = ifelse(present == "", "None", present)) %>%
+  gather(gene, value, -c(present,CIP)) %>%
+  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
+         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
+  na.omit()
+
+p2 <- ggplot(dot_report, aes(reorder(present, -CIP), gene, fill = gene))+
+  geom_point(size = 2, pch = 21)+
+  scale_fill_viridis(option = "B",
+                     discrete = TRUE)+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_text(size = 8),
+        axis.title = element_blank(),
+        axis.ticks.x = element_blank())
+
+plot_grid(p1, p2, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))
+
+```
+
 # Figures
@@
     ) +
-    scale_x_discrete(limits = c(
-      "gyrA", "gyrB", "parC", "parE",
-      "marR", "soxS", "tolC"
-    )) +
     scale_fill_brewer(palette = "Paired") +

### 2018-10-09 notebooks/article_notebook.Rmd
updated tables

@@
 
+## Percent isolates with PMQR
+```{r}
+acquired_report_fixed %>%
+  select(-id) %>%
+  mutate_all(funs(as.numeric(.))) %>%
+  mutate(total = rowSums(.),
+         result = if_else(total != 0, 1, 0)) %>%
+  group_by(result) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
+  spread(result, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr))
+```
+
+## Percent isolates with specific qnr genes
+```{r}
+acquired_report_fixed %>%
+  select(-id) %>%
+  t() %>%
+  as.data.frame(.) %>%
+  mutate(Gene = rownames(.)) %>%
+  mutate_at(vars(-Gene),
+            funs(as.numeric(as.character(.)))) %>%
+  mutate(Present = rowSums(.[,-286])) %>%
+  select(c(Gene, Present)) %>%
+  rowwise() %>%
+  mutate(Total = 285,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+
+```
+
+## Percent isolates with mutations in QRDR
+```{r}
+mut_report_fixed %>%
+  select(-c(id, marR)) %>%
+  group_by_all() %>%
+  count() %>%
+  ungroup() %>%
+  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
+            .funs = funs(as.numeric)) %>%
+  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
+  select(n, type) %>%
+  group_by(type) %>%
+  summarise_all(funs(sum)) %>%
+  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
+  spread(type, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+```
+
+## Percent of isolates with specific mutation combinations in QRDR
+```{r}
+mut_quant_fixed %>%
+  group_by(gyrA, gyrB, parC, parE) %>%
+  count() %>%
+  rowwise() %>%
+  mutate(Total = 285,
+         Percent = round(n/Total * 100,2),
+         lwr = round(get_binCI(n, Total)[1], 2),
+         upr = round(get_binCI(n, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+```
+
+## Total combinations
+```{r}
+acquired_report_fixed %>%
+  left_join(mut_quant_fixed[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
+  left_join(mut_report_fixed[c("marR","id")], by = "id") %>%
+  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
+  select(-id) %>%
+  group_by_at(vars(-c(CIP, NAL))) %>%
+  summarise(CIP = ifelse(min(as.numeric(CIP)) == max(as.numeric(CIP)),
+                         as.character(median(as.numeric(CIP))),
+                         paste(min(as.numeric(CIP)), max(as.numeric(CIP)), sep = " - ")),
+            NAL = ifelse(min(as.numeric(NAL)) == max(as.numeric(NAL)),
+                         as.character(median(as.numeric(NAL))),
+                         paste(min(as.numeric(NAL)), max(as.numeric(NAL)), sep = " - ")),
+            n = n()) %>%
+  rowwise() %>%
+  mutate(Total = 285,
+         Percent = round(n/Total * 100,1),
+         lwr = round(get_binCI(n, Total)[1], 1),
+         upr = round(get_binCI(n, Total)[2], 1),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+```
+
+# Figures
 ## Percent isolates with PMQR and/or QRDR mutations in gyrA, gyrB, parC or parE
@@
 
-# Figures
 ## Percent QREC isolates with mutations in intrinsic genes

### 2018-10-12 notebooks/article_notebook.Rmd
updated with figures on read number densities

--- 2018-10-04 notebooks/article_notebook.Rmd
@@
 
-```{r, include=FALSE}
-id <- c("2017-01-3941-1-S", "2016-17-98-1-S", "2016-17-610-1-S", "2016-17-604-1-S", 
-"2016-17-565-1-S", "2016-17-550-1-S", "2016-17-548-1-S", "2016-17-546-1-S", 
-"2016-17-485-1-S", "2016-17-469-1-S", "2016-17-423-1-S", "2016-17-412-1-S", 
-"2016-17-408-1-T", "2016-17-408-1-S", "2016-17-393-1-S", "2016-17-391-1-S", 
-"2016-17-388-1-T", "2016-17-388-1-S", "2016-17-377-1-S", "2016-17-363-1-S", 
-"2016-17-352-1-S", "2016-17-343-1-S", "2016-17-340-1-S", "2016-17-336-1-S", 
-"2016-17-329-1-S", "2016-17-299-1-S", "2016-17-298-1-S", "2016-17-294-1-S", 
-"2016-17-292-1-S", "2016-17-288-1-S", "2016-17-281-1-T", "2016-17-281-1-S", 
-"2016-17-275-1-S", "2016-17-271-1-S", "2016-17-27-1-S", "2016-17-264-1-S", 
-"2016-17-263-1-S", "2016-17-257-1-S", "2016-17-256-1-S", "2016-17-242-1-S", 
-"2016-17-225-1-S", "2016-17-224-1-S", "2016-17-189-1-T", "2016-17-183-1-S", 
-"2016-17-174-1-T", "2016-17-174-1-S", "2016-17-164-1-S", "2016-17-16-1-S", 
-"2016-17-148-1-S", "2016-17-124-1-S", "2016-17-114-1-S", "2016-17-105-1-T", 
-"2016-17-1-1-S", "2016-02-785-2-S", "2016-02-783-2-T", "2016-02-771-2-S", 
-"2016-02-770-2-S", "2016-02-741-2-S", "2016-02-739-2-S", "2016-02-732-2-S", 
-"2016-02-727-2-S", "2016-02-726-2-S", "2016-02-725-2-S", "2016-02-723-2-S", 
-"2016-02-697-2-S", "2016-02-693-2-T", "2016-02-693-2-S", "2016-02-692-2-S", 
-"2016-02-690-2-T", "2016-02-690-2-S", "2016-02-620-2-S", "2016-02-563-2-S", 
-"2016-02-546-2-S", "2016-02-543-2-S", "2016-02-531-2-S", "2016-02-522-2-S", 
-"2016-02-515-2-T", "2016-02-515-2-S", "2016-02-514-2-S", "2016-02-513-2-T", 
-"2016-02-513-2-S", "2016-02-505-2-S", "2016-02-492-2-S", "2016-02-490-2-S", 
-"2016-02-486-2-S", "2016-02-447-2-S", "2016-02-446-2-S", "2016-02-428-2-S", 
-"2016-02-425-2-S", "2016-02-419-2-S", "2016-02-418-2-S", "2016-02-415-2-S", 
-"2016-02-413-2-T", "2016-02-413-2-S", "2016-02-411-2-S", "2016-02-410-2-S", 
-"2016-02-409-2-S", "2016-02-408-2-S", "2016-02-402-2-S", "2016-02-381-2-S", 
-"2016-02-369-2-S", "2016-02-366-2-S", "2016-02-363-2-S", "2016-02-359-2-S", 
-"2016-02-358-2-S", "2016-02-356-2-S", "2016-02-353-2-S", "2016-02-352-2-S", 
-"2016-02-351-2-S", "2016-02-348-2-S", "2016-02-345-2-T", "2016-02-345-2-S", 
-"2016-02-343-2-S", "2016-02-336-2-S", "2016-02-330-2-S", "2016-02-329-2-S", 
-"2016-02-324-2-S", "2016-02-316-2-S", "2016-02-302-2-S", "2016-02-293-2-S", 
-"2016-02-281-2-S", "2016-02-278-2-S", "2016-01-3763-1-T", "2016-01-3474-1-T", 
-"2015-01-753-1-S", "2015-01-752-1-S", "2015-01-751-1-S", "2015-01-750-1-S", 
-"2015-01-749-1-S", "2015-01-5263-1-S", "2015-01-5262-1-S", "2015-01-5177-1-S", 
-"2015-01-5141-1-S", "2015-01-5140-1-S", "2015-01-5138-1-S", "2015-01-5022-1-S", 
-"2015-01-5020-1-S", "2015-01-5018-1-S", "2015-01-4988-1-S", "2015-01-4934-1-S", 
-"2015-01-4932-1-S", "2015-01-4931-1-S", "2015-01-4794-1-S", "2015-01-4788-1-S", 
-"2015-01-4742-1-S", "2015-01-467-1-S", "2015-01-466-1-S", "2015-01-463-1-S", 
-"2015-01-4559-1-S", "2015-01-4558-1-S", "2015-01-4465-1-S", "2015-01-4463-1-S", 
-"2015-01-4462-1-S", "2015-01-4421-1-S", "2015-01-4249-1-S", "2015-01-4037-1-S", 
-"2015-01-3915-1-S", "2015-01-3914-1-S", "2015-01-3911-1-S", "2015-01-3887-1-S", 
-"2015-01-3640-1-S", "2015-01-3639-1-S", "2015-01-3638-1-S", "2015-01-3548-1-S", 
-"2015-01-3545-1-S", "2015-01-3318-1-S", "2015-01-3317-1-S", "2015-01-3102-1-S", 
-"2015-01-3004-1-S", "2015-01-3001-1-S", "2015-01-2879-1-S", "2015-01-2878-1-S", 
-"2015-01-2839-1-S", "2015-01-2838-1-S", "2015-01-2837-1-S", "2015-01-2836-1-S", 
-"2015-01-2835-1-S", "2015-01-2771-1-S", "2015-01-2430-1-S", "2015-01-2429-1-S", 
-"2015-01-2426-1-S", "2015-01-2404-1-S", "2015-01-2403-1-S", "2015-01-2309-1-S", 
-"2015-01-2166-1-S", "2015-01-2097-1-S", "2015-01-2094-1-S", "2015-01-1955-1-S", 
-"2015-01-1954-1-S", "2015-01-1952-1-T", "2015-01-1951-1-S", "2015-01-1950-1-S", 
-"2015-01-1949-1-S", "2015-01-1826-1-S", "2015-01-1823-1-S", "2015-01-1661-1-S", 
-"2015-01-1566-1-S", "2015-01-1343-1-S", "2014-01-861-1-S", "2014-01-860-1-S", 
-"2014-01-856-1-S", "2014-01-7584-1-S", "2014-01-7543-1-S", "2014-01-7512-1-S", 
-"2014-01-7490-1-S", "2014-01-7479-1-S", "2014-01-7443-1-S", "2014-01-7442-1-S", 
-"2014-01-7394-1-S", "2014-01-7375-1-S", "2014-01-7234-1-S", "2014-01-7207-1-S", 
-"2014-01-7136-1-S", "2014-01-7133-1-S", "2014-01-7050-1-S", "2014-01-6926-1-S", 
-"2014-01-6924-1-S", "2014-01-6710-1-S", "2014-01-6625-1-S", "2014-01-6530-1-S", 
-"2014-01-6291-1-S", "2014-01-6123-1-S", "2014-01-6122-1-S", "2014-01-6044-1-S", 
-"2014-01-6043-1-S", "2014-01-6041-1-S", "2014-01-6040-1-S", "2014-01-5917-1-T", 
-"2014-01-5917-1-S", "2014-01-5916-1-S", "2014-01-5915-1-S", "2014-01-5912-1-S", 
-"2014-01-5792-1-S", "2014-01-5565-1-S", "2014-01-4922-1-T", "2014-01-4921-1-S", 
-"2014-01-4651-1-S", "2014-01-4539-1-S", "2014-01-4271-1-S", "2014-01-4270-1-S", 
-"2014-01-3802-1-S", "2014-01-3677-1-S", "2014-01-3676-1-S", "2014-01-3212-1-S", 
-"2014-01-2996-1-S", "2014-01-2995-1-S", "2014-01-2992-1-S", "2014-01-2773-1-S", 
-"2014-01-2728-1-S", "2014-01-2516-1-S", "2014-01-2366-1-S", "2014-01-2362-1-T", 
-"2014-01-2145-1-S", "2014-01-2144-1-S", "2014-01-2071-1-S", "2014-01-2070-1-S", 
-"2014-01-2069-1-S", "2014-01-1742-1-S", "2014-01-1741-1-S", "2014-01-1649-1-S", 
-"2014-01-1268-1-T", "2014-01-1182-1-S", "2013-01-2024-1-T", "2012-01-838-6-T", 
-"2011-01-9386-3-T", "2011-01-6927-5-T", "2011-01-3460-5-T", "2011-01-1544-2-T", 
-"2010-01-991-7-T", "2009-01-551-1-T", "2009-01-4672-5-T", "2009-01-4618-2-T", 
-"2009-01-4415-4-T", "2009-01-413-1-T", "2009-01-4064-1-T", "2009-01-3815-1-T", 
-"2009-01-3064-2-T", "2009-01-2696-4-T", "2009-01-2521-3-T", "2009-01-175-4-T", 
-"2009-01-175-2-T", "2009-01-1690-1-T", "2007-01-361-1-T", "2006-01-2821-1-T", 
-"2006-01-2184-1-T")
-```
+```{r Import data, include = FALSE}
 
-```{r Import data, include = FALSE}
+# Collapses data frame to unique lines while ignoring NA's
+func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")
+
+id_data <- read.table("../data/id.txt",
+                      sep = "\t",
+                      header = TRUE,
+                      stringsAsFactors = F)
+
 isolate_data <- read.table("../data/isolate_data.txt",
@@
 
-no_of_reads <- read.table("../data/read_data.txt",
+two_reads <- read.table("../data/no_of_reads.txt",
+                          sep = "\t",
+                          header = F,
+                          stringsAsFactors = F) %>%
+  rename("ref" = V1,
+         "reads" = V2) %>%
+  mutate(ref = gsub(".+/(.*?)\\.fastq\\.gz", "\\1", ref),
+         type = ifelse(grepl("R1", ref) == TRUE, "R1", "R2")) %>%
+  mutate(ref = gsub("(.*?)_L00[0-9]_R[1-2]_00[0-9]","\\1", ref)) %>%
+  group_by(ref) %>%
+  mutate(id = n()) %>%
+  {. ->> four_reads} %>%
+  filter(id == 2) %>%
+  spread(type, reads) %>%
+  rename("R1-1" = R1,
+         "R2-1" = R2) %>%
+  mutate("R1-2" = NA,
+         "R2-2" = NA) %>%
+  select(-id) %>%
+  mutate_all(funs(as.character(.)))
+
+four_reads <- four_reads %>%
+  filter(id == 4) %>%
+  group_by(ref) %>%
+  mutate(id = 1:n()) %>%
+  spread(id, reads) %>%
+  summarise_all(funs(func_paste)) %>%
+  rename("R1-1" = `1`,
+         "R2-1" = `2`,
+         "R1-2" = `3`,
+         "R2-2" = `4`) %>%
+  select(-type) %>%
+  mutate_all(funs(as.character(.)))
+
+no_of_reads <- as.data.frame(rbind(as.matrix(two_reads), as.matrix(four_reads))) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id, everything(), -ref)
+
+mic_data <- read.table("../data/mic_data.txt",
                            sep = "\t",
@@
 
+mlst_hf150bp <- read.table("../data/mlst/hiseq_flex_150bp_mlst_summarized_results.tsv",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
+
+mlst_hx125bp <- read.table("../data/mlst/hiseq_xt_125bp_mlst_summarized_results.tsv",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
+
+mlst_hrx250bp <- read.table("../data/mlst/hiseqr_xt_250bp_mlst_summarized_results.tsv",
+                           sep = "\t",
+                           header = T,
+                           stringsAsFactors = F)
+
+mlst_data <- rbind(mlst_hf150bp, mlst_hx125bp, mlst_hrx250bp) %>%
+  separate(ST.adk.fumC.gyrB.icd.mdh.purA.recA,
+           into = c("x","ST","adk","fumC","gyrB","icd","mdh","purA","recA"),
+           sep = " ") %>%
+  select(-x) %>%
+  rename("ref" = header) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id, everything(), -ref)
+```
+
+# Sequence data
+## Distribution of number of reads per sequencer
+```{r}
+sequence_data <- no_of_reads %>%
+  left_join(isolate_data, by = "id") %>%
+  mutate_at(vars(`R1-1`,`R2-1`,`R1-2`,`R2-2`),
+            funs(as.numeric(as.character(.))))
+
+ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
+  geom_density(aes(fill = sequencer,
+                   y = ..scaled..),
+               alpha = 0.4)+
+  labs(x = "Number of reads",
+       y = "Density",
+       fill = "Sequencer")+
+  theme(legend.position = c(0.7, 0.8))
+```
+
+## Distribution of number of reads per library prep type
+```{r}
+ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
+  geom_density(aes(fill = libprep,
+                   y = ..scaled..),
+               alpha = 0.4)+
+  labs(x = "Number of reads",
+       y = "Density",
+       fill = "Library prep kit")+
+  theme(legend.position = c(0.7, 0.8))
+```
+
+## Distribution of number of reads per animal species
+```{r}
+ggplot(sequence_data, aes(as.numeric(`R1-1`)))+
+  geom_density(aes(fill = species,
+                   y = ..scaled..),
+               alpha = 0.4)+
+  labs(x = "Number of reads",
+       y = "Density",
+       fill = "Animal species")+
+  theme(legend.position = c(0.7, 0.8))
 ```
@@
 # -------------------------------------- Functions
-# Collapses data frame to unique lines while ignoring NA's
-func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")
+
+func_paste2 <- function(x) paste(unique(x[!is.na(x)]), collapse = ",")
 
@@
     mutate(flag_result = flag %in% flag_selection,
-           flag_result = as.integer(flag_result),
-           ref = gsub("(.*?)\\_.+", "\\1", ref),
-           ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
-           ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)) %>%
+           flag_result = as.integer(flag_result)) %>%
     rename("gene" = gene_names)
@@
   df <- df %>%
-    select(ref, gene_names, flag, ref_ctg_change) %>%
+    mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
+           ref_nt = ifelse(ref_nt == ".", "", ref_nt),
+           ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
+           ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
+           mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
+           mutation = ifelse(mutation == "/-", "0", mutation)) %>%
+    select(ref, gene_names, flag, mutation) %>%
     filter(flag %in% flag_selection) %>%
     mutate(id = 1:n()) %>%
-    spread(gene_names, ref_ctg_change) %>%
+    spread(gene_names, mutation) %>%
     group_by(ref) %>%
@@
            result = as.integer(result),
-           type = "mut")
+           type = "mut",
+           nt_change = gsub("[^,]+?/([^,]+?)", "\\1", mut),
+           mut = gsub("([^/]+)/[^,]+", "\\1", mut))
   return(df)
@@
     select(-mut) %>%
-    group_by(ref) %>%
-    mutate(id = 1:n()) %>%
+    group_by(id) %>%
+    mutate(id2 = 1:n()) %>%
     spread(gene, result_total) %>%
     summarise_all(funs(func_paste)) %>%
-    select(-c(id, type))
+    select(-c(id2, type))
   return(df)
@@
   df <- df %>%
-    group_by(ref) %>%
-    mutate(id = 1:n()) %>%
+    group_by(id) %>%
+    mutate(id2 = 1:n()) %>%
     spread(gene, result_total) %>%
     summarise_all(funs(func_paste)) %>%
-    select(-c(id, type))
+    select(-c(id2, type))
   return(df)
-}
-
-# prepares data frame for number of mutations heatmap
-create_heatmap_df <- function(df) {
-  qnr <- names(df)
-
-  qnr_cols <- c()
-
-  for (i in qnr) {
-    if (grepl("qnr", i, ignore.case = TRUE) == TRUE) {
-      qnr_cols <- c(qnr_cols, i)
-    }
-  }
-
-  qnr_df <- df %>%
-    select_at(vars(ref, qnr_cols)) %>%
-    gather(gene, result, -ref) %>%
-    mutate(type = "Gene")
-
-  qnr_mut_quant <- mut_quant %>%
-    select(-c("gyrA","gyrB","parC","parE")) %>%
-    mutate(type = "Mut") %>%
-    gather(gene, result, -c(ref, type)) %>%
-    mutate(gene = case_when(gene == "mut_gyrA" ~ "gyrA",
-                            gene == "mut_gyrB" ~ "gyrB",
-                            gene == "mut_parC" ~ "parC",
-                            gene == "mut_parE" ~ "parE",
-                            TRUE ~ gene)) %>%
-    rbind(., qnr_df)
-  return(qnr_mut_quant)
 }
@@
 # Check flags
-mut_flags <- check_flags(clean_mut_data)
-acquired_flags <- check_flags(clean_acquired_data)
+mut_flags <- check_flags(clean_mut_data) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id, everything(), -ref)
+
+acquired_flags <- check_flags(clean_acquired_data) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id, everything(), -ref)
 
@@
 mut_table <- create_mut_table(clean_mut_data)
-mut_table_fixed <- fix_gyr_par_results(mut_table)
+mut_table_fixed <- fix_gyr_par_results(mut_table) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id, everything(), -ref)
+
 mut_filtered <- filter_mut_table(mut_table_fixed)
-acquired_table <- create_acquired_table(clean_acquired_data)
+acquired_table <- create_acquired_table(clean_acquired_data) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id, everything(), -ref)
+
 acquired_filtered <- filter_acquired_table(acquired_table)
@@
 
-mut_quant <- suppressWarnings(calc_no_of_mut(mut_table))
-heatmap_df <- create_heatmap_df(acquired_report)
+mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
+  left_join(id_data, by = "ref") %>%
+  select(id, everything(), -ref)
 
@@
 
+## Percent isolates with PMQR
+```{r}
+acquired_report %>%
+  select(-id) %>%
+  mutate_all(funs(as.numeric(.))) %>%
+  mutate(total = rowSums(.),
+         result = if_else(total != 0, 1, 0)) %>%
+  group_by(result) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(result = if_else(result == 0, "Absent", "Present")) %>%
+  spread(result, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr))
+```
+
+## Percent isolates with specific qnr genes
+```{r}
+acquired_report %>%
+  select(-id) %>%
+  t() %>%
+  as.data.frame(.) %>%
+  mutate(Gene = rownames(.)) %>%
+  mutate_at(vars(-Gene),
+            funs(as.numeric(as.character(.)))) %>%
+  mutate(Present = rowSums(.[,-286])) %>%
+  select(c(Gene, Present)) %>%
+  rowwise() %>%
+  mutate(Total = 285,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+```
+
+## Percent isolates with mutations in QRDR
+```{r}
+mut_report %>%
+  select(-c(id, marR, nt_change)) %>%
+  group_by_all() %>%
+  count() %>%
+  ungroup() %>%
+  mutate_at(vars(c(gyrA, gyrB, parC, parE)),
+            .funs = funs(as.numeric)) %>%
+  mutate(type = ifelse(rowSums(.[,1:4]) == 0, 1, 0)) %>%
+  select(n, type) %>%
+  group_by(type) %>%
+  summarise_all(funs(sum)) %>%
+  mutate(type = ifelse(type == 0, "Present", "Absent")) %>%
+  spread(type, n) %>%
+  mutate(Total = Present + Absent,
+         Percent = round(Present/Total * 100,2),
+         lwr = round(get_binCI(Present, Total)[1], 2),
+         upr = round(get_binCI(Present, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+```
+
+## Percent of isolates with specific mutation combinations in QRDR
+```{r}
+mut_quant %>%
+  group_by(gyrA, gyrB, parC, parE) %>%
+  count() %>%
+  rowwise() %>%
+  mutate(Total = 285,
+         Percent = round(n/Total * 100,2),
+         lwr = round(get_binCI(n, Total)[1], 2),
+         upr = round(get_binCI(n, Total)[2], 2),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+```
+
+## Percent isolates with specific nucleotide changes in QRDR
+```{r}
+suppressWarnings(clean_mut_data %>%
+  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
+         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
+         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
+         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
+         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
+         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
+  select(ref, gene_names, flag, mutation) %>%
+  filter(flag %in% flag_selection,
+         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
+  mutate(id = 1:n()) %>%
+  spread(gene_names, mutation) %>%
+  group_by(ref) %>%
+  summarise_all(funs(func_paste2)) %>%
+  select(-c(id, flag)) %>%
+  gather(gene, mut, -ref) %>%
+  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
+         result = ifelse(mut != 0, 1, 0),
+         result = as.integer(result),
+         type = "mut") %>%
+  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
+  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
+  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
+  select(-mut) %>%
+  rename("mut" = aa_change) %>%
+  filter(is.na(nt_change) == FALSE) %>%
+  fix_gyr_par_results(.) %>%
+  filter(result_total == 1) %>%
+  group_by(gene, mut, nt_change) %>%
+  count() %>%
+  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
+  rowwise() %>%
+  mutate(Total = 285,
+         Percent = round(n/Total * 100,1),
+         lwr = round(get_binCI(n, Total)[1],1),
+         upr = round(get_binCI(n, Total)[2],1),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total)) %>%
+  rename("Mutation" = mut))
+```
+
+## Total combinations
+```{r}
+test <- acquired_report %>%
+  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
+  left_join(mut_report[c("marR","id")], by = "id") %>%
+  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
+  select(-id) %>%
+  group_by_at(vars(-c(CIP, NAL))) %>%
+  summarise(CIP = ifelse(min(as.numeric(CIP)) == max(as.numeric(CIP)),
+                         as.character(median(as.numeric(CIP))),
+                         paste(min(as.numeric(CIP)), max(as.numeric(CIP)), sep = " - ")),
+            NAL = ifelse(min(as.numeric(NAL)) == max(as.numeric(NAL)),
+                         as.character(median(as.numeric(NAL))),
+                         paste(min(as.numeric(NAL)), max(as.numeric(NAL)), sep = " - ")),
+            n = n()) %>%
+  rowwise() %>%
+  mutate(Total = 285,
+         Percent = round(n/Total * 100,1),
+         lwr = round(get_binCI(n, Total)[1], 1),
+         upr = round(get_binCI(n, Total)[2], 1),
+         "95 % CI" = paste("[", lwr, "-", upr, "]")) %>%
+  select(-c(lwr, upr, Total))
+```
+
 # Figures
+## Percent isolates with PMQR and/or QRDR mutations in gyrA, gyrB, parC or parE
+```{r}
+total_report <- mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  gather(gene, value, -id) %>%
+  mutate(present = ifelse(value == 1, gene, NA),
+         ref = 1:n()) %>%
+  spread(gene, value) %>%
+  group_by(id) %>%
+  summarise_all(funs(func_paste)) %>%
+  select(-ref) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  group_by(present,CIP) %>%
+  count() %>%
+  ungroup() %>%
+  rowwise() %>%
+  mutate(present = ifelse(present == "", "None", present),
+         total = 285,
+         percent = round(n/total * 100, 1),
+         lwr = round(get_binCI(n, total)[1], 1),
+         upr = round(get_binCI(n, total)[2], 1))
+
+genes <- c(names(mut_report), names(acquired_report))
+genes <- genes[!genes == "id"]
+
+p1 <- ggplot(total_report, aes(reorder(present, as.numeric(-CIP)), percent, fill = factor(CIP)))+
+  geom_col(color = "black")+
+  labs(y = "Percent (%) isolates")+
+  scale_fill_viridis(discrete = TRUE)+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_text(size = 10),
+        axis.text.y = element_text(size = 8),
+        axis.title.x = element_blank(),
+        legend.text = element_text(size = 8),
+        legend.title = element_blank(),
+        legend.position = "top",
+        legend.spacing.x = unit(0.27, "cm"))
+
+dot_report <- mut_report %>%
+  left_join(acquired_report, by = "id") %>%
+  gather(gene, value, -id) %>%
+  mutate(present = ifelse(value == 1, gene, NA),
+         ref = 1:n()) %>%
+  spread(gene, value) %>%
+  group_by(id) %>%
+  summarise_all(funs(func_paste)) %>%
+  left_join(mic_data[c("id","CIP")], by = "id") %>%
+  select(-c(ref, id)) %>%
+  ungroup() %>%
+  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
+  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
+         present = ifelse(present == "", "None", present)) %>%
+  gather(gene, value, -c(present,CIP)) %>%
+  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
+         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
+  na.omit()
+
+p2 <- ggplot(dot_report, aes(reorder(present, -CIP), gene, fill = gene))+
+  geom_point(size = 2, pch = 21)+
+  scale_fill_viridis(option = "B",
+                     discrete = TRUE)+
+  guides(fill = FALSE)+
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_text(size = 8),
+        axis.title = element_blank(),
+        axis.ticks.x = element_blank())
+
+plot_grid(p1, p2, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))
+
+```
+
 ## Percent QREC isolates with mutations in intrinsic genes
@@
 
-gene_names <- gene_names[-1]
+gene_names <- gene_names[-c(1,2)]
 
 mut_report %>%
-    mutate(
-      ref = gsub("(.*?)\\_.+", "\\1", ref),
-      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
-      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
-    ) %>%
-    arrange(desc(ref)) %>%
-    mutate(id = id) %>%
     left_join(isolate_data, by = "id") %>%
@@
 acquired_report %>%
-    mutate(
-      ref = gsub("(.*?)\\_.+", "\\1", ref),
-      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
-      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
-    ) %>%
-    arrange(desc(ref)) %>%
-    mutate(id = id) %>%
     left_join(isolate_data, by = "id") %>%
